---
title: "Projected Health Impact of WHO Cardiovascular Risk-Factor Targets, 2026--2050"
author: ""
institute: "WHO $\\cdot$ University of Washington"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    theme: "default"
    keep_tex: false
    latex_engine: xelatex
    slide_level: 2
    toc: true
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
---

<!-- ======================================================================
     SETUP — LOAD ARTEFACTS ONLY (zero recomputation)
     All objects computed by aim1_report.Rmd and saved to output/slides/
     ====================================================================== -->

```{r load-artefacts, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)

# ---- Path to pre-built artefacts --------------------------------------------
wd_outp    <- "C:/Users/wrgar/OneDrive - UW/02Work/WHO-CVD/github/uw-who-cvdtargets/output/"
wd_data    <- "C:/Users/wrgar/OneDrive - UW/02Work/WHO-CVD/github/uw-who-cvdtargets/data/processed/"
slides_dir <- paste0(wd_outp, "slides/")

stopifnot(
  "output/slides/ not found — knit aim1_report.Rmd first." =
    dir.exists(slides_dir),
  "sl_scalars.rds missing — knit aim1_report.Rmd first." =
    file.exists(paste0(slides_dir, "sl_scalars.rds"))
)

# ---- SL-01: Scalars ----------------------------------------------------------
sl <- readRDS(paste0(slides_dir, "sl_scalars.rds"))
n_countries      <- sl$n_countries
study_years      <- sl$study_years
total_all        <- sl$total_all
total_bp         <- sl$total_bp
total_bpd        <- sl$total_bpd
total_stat       <- sl$total_stat
bau_total_deaths <- sl$bau_total_deaths

# ---- SL-02 to SL-08: Data tables --------------------------------------------
sl_tab_summary <- readRDS(paste0(slides_dir, "sl_tab_summary.rds"))
sl_tab_region  <- readRDS(paste0(slides_dir, "sl_tab_region.rds"))
sl_tab_top15   <- readRDS(paste0(slides_dir, "sl_tab_top15.rds"))
sl_dt_cause    <- readRDS(paste0(slides_dir, "sl_dt_cause.rds"))
sl_dt_age      <- readRDS(paste0(slides_dir, "sl_dt_age.rds"))
sl_dt_asmr     <- readRDS(paste0(slides_dir, "sl_dt_asmr.rds"))
sl_dt_annual   <- readRDS(paste0(slides_dir, "sl_dt_annual.rds"))

# ---- SL-09 to SL-14: PNG paths ----------------------------------------------
png_trend       <- paste0(slides_dir, "sl_fig_baseline_trend.png")
png_asmr        <- paste0(slides_dir, "sl_fig_asmr.png")
png_averted     <- paste0(slides_dir, "sl_fig_deaths_averted.png")
png_cum_averted <- paste0(slides_dir, "sl_fig_cum_averted.png")
png_cause       <- paste0(slides_dir, "sl_fig_by_cause.png")
png_age         <- paste0(slides_dir, "sl_fig_by_age.png")
png_top15       <- paste0(slides_dir, "sl_fig_top15_countries.png")

# ---- Display helpers (no raw data needed) ------------------------------------
fmt_M <- function(x, digits = 1)
  formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")

int_order <- c(
  "HTN Control (General)",
  "HTN Control (Diabetes)",
  "Improved Statin Uptake",
  "All Interventions"
)

int_cols <- c(
  "HTN Control (General)"  = "#1f78b4",
  "HTN Control (Diabetes)" = "#a6cee3",
  "Improved Statin Uptake" = "#e31a1c",
  "All Interventions"      = "#6a3d9a",
  "Business-as-Usual"      = "#999999"
)

knitr::opts_chunk$set(
  echo      = FALSE,
  message   = FALSE,
  warning   = FALSE,
  fig.align = "center",
  fig.height = 3.6,
  fig.width  = 8.5,
  out.width  = "\\linewidth",
  dev = "cairo_pdf"
)
```

<!-- ======================================================================
     TITLE / CONTEXT
     ====================================================================== -->

# Study Overview

## Aim 1: Research Questions

\begin{block}{Central Question}
What is the projected \textbf{burden-of-disease impact} of achieving the WHO cardiovascular risk-factor control targets by \textbf{2030}?
\end{block}

\textbf{Sub-questions driving the analysis:}

\begin{enumerate}
  \item How many \textbf{cardiovascular events} could be averted?
  
  \item How many \textbf{CVD deaths} could be prevented by 2030 and beyond?
  
  \item How would \textbf{life-years, YLLs, YLDs, and DALYs} change under joint achievement of:
  \begin{itemize}
      \item 150M additional people with controlled hypertension
      \item $\geq$50\% lipid-lowering therapy coverage
      \item 80\% BP control among people with diabetes
      \item \alert{80\% validated BP device availability}
  \end{itemize}
\end{enumerate}

\medskip
\begin{columns}[T]

\begin{column}{0.48\textwidth}
\begin{exampleblock}{Scope}
\small
\textbf{`r n_countries` countries} $\cdot$ Ages 20--95 $\cdot$ 2026--2050 $\cdot$ Multiple CVD causes $\cdot$ WHO Target Scenarios
\end{exampleblock}
\end{column}

\begin{column}{0.48\textwidth}
\begin{alertblock}{Intervention start: 2026}
\small
Model calibrated to 2019 baseline. First year of target scale-up effects entering the projection is \textbf{2026}.
\end{alertblock}
\end{column}

\end{columns}

## Study Design at a Glance

```{r design-table}
design <- data.table(
  Component = c("Model type","Calibration data","Population data",
                "Intervention start","Scale-up period","Comparison",
                "Outcome","Time horizon"),
  Detail = c(
    "Discrete-time multi-state model (Well $\\to$ Sick $\\to$ Dead)",
    "GBD 2023: incidence, case fatality, prevalence; all-cause \\& CVD-specific",
    "UNWPP 2024 single-year age/sex projections",
    "\\textbf{2026} --- first year intervention effects enter the projection",
    "Linear scale-up 2026 $\\to$ 2030; rates held constant 2030--2050",
    "Business-as-usual: 2024 control rates, no additional scale-up",
    "CVD deaths averted vs.~BAU; age-standardised mortality rate (ASMR)",
    "2026--2050 (25-year horizon); model calibrated from 2019"
  )
)
kable(design, col.names=c("Component","Detail"),
      align="ll", escape=FALSE, booktabs=TRUE, format="latex") |>
  kable_styling(latex_options=c("striped","hold_position"),
                font_size=8, full_width=TRUE) |>
  column_spec(1, bold=TRUE, width="3.2cm") |>
  column_spec(2, width="9.8cm") |>
  row_spec(4, background="#dbeafe")
```

## Study Context

\begin{columns}[T]
\begin{column}{0.55\textwidth}
\textbf{Objective}
\begin{itemize}
  \item Quantify CVD deaths avertable by achieving WHO 2030 cardiovascular targets
  \item Coverage: \textbf{`r n_countries` countries}, `r study_years`
  \item Four causes: IHD, ischemic stroke, hemorrhagic stroke, hypertensive HD
\end{itemize}
\vspace{6pt}
\textbf{Model}
\begin{itemize}
  \item Discrete-time state transition (Well $\to$ Sick $\to$ Dead)
  \item Calibrated to GBD 2023; projected with UNWPP 2024 populations
  \item 8 blood-pressure bins
  \item Baseline treatment effects explicitly de-duplicated
\end{itemize}
\end{column}
\begin{column}{0.43\textwidth}
\begin{exampleblock}{Research Question}
What are the projected deaths averted from achieving WHO hypertension control and lipid-lowering therapy targets across \textbf{`r n_countries` countries}, 2026--2050?
\end{exampleblock}
\begin{block}{Pending \textcolor{red!70!black}{$\bigtriangleup$}}
Confirm primary metric: deaths averted vs.\ DALYs vs.\ LYG
\end{block}
\end{column}
\end{columns}

# Data Sources

## Data Sources and Key Variables

```{r data-sources-table}
ds <- data.table(
  Source    = c("GBD 2023 (IHME)","GBD 2019 (IHME)","Ettehad et al.\\ 2016",
                "UNWPP 2024 (UN)","WHO / NCD-RisC","Country-specific"),
  Variables = c(
    "IR, CF, prevalence, background mortality (age/sex/country)",
    "Relative risks per 10 mmHg SBP increase (IHD, stroke, HHD, by age)",
    "Trial effect sizes per BP category; diabetes-stratified",
    "Population projections 2019--2050 (single-year age, sex, country)",
    "Baseline HTN control rates ($C_0$); BP distribution by country",
    "HTN control targets 2030 (Progress / Ambitious / Aspirational)"
  ),
  Role = c(
    "Baseline transition rates; initial states",
    "Anchor BP-bin incidence; normalisation factor $\\alpha$",
    "Treatment effect sizes by BP bin $\\times$ cause",
    "Cohort sizes; rate denominators",
    "$C_0$; $P(\\text{BP}_{\\text{cat}})$",
    "$C_{\\text{target}}$ per scenario"
  )
)
kable(ds, align="lll", escape=FALSE, booktabs=TRUE, format="latex",
      col.names=c("Source","Variables extracted","Role in model")) |>
  kable_styling(latex_options=c("striped","hold_position"),
                font_size=8, full_width=TRUE) |>
  column_spec(1, bold=TRUE, width="2.8cm") |>
  column_spec(2, width="6.8cm") |>
  column_spec(3, width="3.4cm")
```

<!-- ======================================================================
     SCENARIOS
     ====================================================================== -->

# Scenarios

## Four WHO Priority Targets

```{r scenario-table, results='asis'}
dt_scen <- data.table(
  N  = 1:4,
  WT = c("150M more with HTN controlled by 2030",
         "$\\geq$50\\% eligible adults on lipid-lowering therapy by 2030",
         "80\\% BP control among diagnosed diabetics by 2030",
         "80\\% availability of validated BP devices"),
  Sc = c("\\textbf{HTN Control (General)}: $\\approx$ 100M more people with BP control by 2030",
         "\\textbf{Improved Statin Uptake}: 50\\% coverage by 2030",
         "\\textbf{HTN Control (Diabetes)}: 80\\% BP control in diabetics by 2030 ($\\approx$ 50M more people with BP control)",
         "Operationally embedded in HTN scenarios (150M Target [1+3])")
)
kable(dt_scen, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("\\#", "WHO Target", "Modelled Scenario"), caption = NULL) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 7) |>
  column_spec(2, width = "4.5cm") |>
  column_spec(3, width = "4.5cm")
```

\vspace{4pt}
\textbf{Combination rule:} $\varepsilon_{IR,\text{total}} = \varepsilon_{IR,\text{BP}} \times \varepsilon_{IR,\text{statin}}$

## HTN Control Targets by 2030 --- Country Distribution

```{r scenario-targets-plot, fig.height=2.9}
dt_hbp_targets <- fread(paste0(wd_data,"htn_control_targets_by_loc.csv"))
dt_long <- melt(
  dt_hbp_targets,
  id.vars = "location",
  measure.vars = c("htncov2", "htncov2_progress"),
  variable.name = "target_col",
  value.name   = "target_value"
)

# Clean scenario labels
dt_long[, scenario_lbl := factor(
  ifelse(target_col == "htncov2", "Current",
         "Progress"),
  levels = c("Current","Progress")
)]

ggplot(dt_long,
       aes(x = scenario_lbl,
           y = target_value*100,
           fill = scenario_lbl)) +
  
  geom_boxplot(alpha=0.75, outlier.size=1, width=0.45) +
  
  annotate("rect",
           xmin=1.5, xmax=2.5,
           ymin=0, ymax=100,
           fill="#2171b5",
           alpha=0.07) +
  
  annotate("text",
           x=2, y=97,
           label="Headline scenario",
           size=3,
           color="#2171b5",
           fontface="bold") +
  
  scale_y_continuous(limits=c(0,100), expand=c(0,0)) +
  
  labs(
    title="Distribution of Country-Level HTN Control Targets",
    subtitle="Current vs 150M scenario | scale-up starts 2026",
    x=NULL,
    y="HTN Control Rate (%)"
  ) +
  
  theme_minimal(base_size=11) +
  theme(
    plot.title = element_text(face="bold"),
    panel.grid.minor = element_blank()
  )
```

<!-- ======================================================================
     MODEL
     ====================================================================== -->
     
# Mathematical Model

## Model Architecture: Three Health States

```{r state-diagram, fig.height=2.7}
nodes <- data.table(
  x=c(1,3.2,5.4), y=c(1,1,1),
  label=c("Well\n(W)","Sick\n(K)","Dead\n(D)"),
  fill=c("#d4edda","#fff3cd","#f8d7da")
)
arr_main <- data.table(
  x1=c(1.40,3.60), y1=c(1,1), x2=c(2.80,5.00), y2=c(1,1),
  lbl=c("IR \u00d7 \u03b5_IR","CF \u00d7 \u03b5_CF")
)
arr_bg <- data.table(
  x1=c(1.0,3.2,3.0), y1=c(1.25,1.25,1.0),
  x2=c(1.0,3.2,5.0), y2=c(0.62,0.62,0.62),
  lbl=c("BG.mx","BG.mx","BG.mx.all")
)
ggplot() +
  geom_tile(data=nodes, aes(x=x,y=y,fill=fill),
            width=0.75,height=0.42,color="grey30",linewidth=0.8) +
  scale_fill_identity() +
  geom_text(data=nodes, aes(x=x,y=y,label=label), size=4, fontface="bold") +
  geom_segment(data=arr_main,
    aes(x=x1,y=y1,xend=x2,yend=y2),
    arrow=arrow(length=unit(0.22,"cm"),type="closed"),
    linewidth=1.2, color="#2171b5") +
  geom_label(data=arr_main,
    aes(x=(x1+x2)/2, y=y1+0.22, label=lbl),
    size=3.2, fill="white", label.padding=unit(0.12,"cm"), color="#2171b5") +
  geom_segment(data=arr_bg,
    aes(x=x1,y=y1,xend=x2,yend=y2),
    arrow=arrow(length=unit(0.16,"cm"),type="closed"),
    linewidth=0.7, color="grey55", linetype="dashed") +
  geom_label(data=arr_bg,
    aes(x=(x1+x2)/2+0.2, y=(y1+y2)/2, label=lbl),
    size=2.6, fill="white", color="grey50", label.padding=unit(0.08,"cm")) +
  scale_x_continuous(limits=c(0.5,6.0)) +
  scale_y_continuous(limits=c(0.28,1.58)) +
  theme_void(base_size=11) +
  labs(title="Multi-State Transition Model  |  Annual time step  |  Ages 20-95  |  4 CVD causes")
```

\small \textbf{Key:} $\varepsilon_{IR}$, $\varepsilon_{CF}$ = multiplicative effect ratios from the intervention module $\cdot$ BG.mx = background mortality $\cdot$ COVID-19 excess mortality applied 2020--2021 only


## Modeling Assumptions

\begin{columns}[T]
\begin{column}{0.47\textwidth}
\textbf{Model structure}
\begin{itemize}\footnotesize
  \item Discrete-time annual step (2019--2050); ages 20--95
  \item Four causes modeled independently
  \item COVID-19 excess mortality applied 2020--2021 only; no long-term effects
\end{itemize}

\medskip
\textbf{Intervention mechanics}
\begin{itemize}\footnotesize
  \item Linear scale-up 2026 $\to$ 2030; held constant 2030--2050
  \item Only incremental coverage above $C_0$ (or $U_{0,\ell}$) generates new effects (no double-counting)
  \item Full adherence assumed for HTN control, statins adherence adjusted down for primary and secondary prevention.
  \item Statin effects restricted to ages $\geq 40$ and IHD \& ischemic stroke only
\end{itemize}
\end{column}

\begin{column}{0.52\textwidth}
\textbf{Blood pressure \& effect sizes}
\begin{itemize}\footnotesize
  \item Static log-normal BP distribution: no secular trend, no endogenous updating as coverage rises
  \item Treatment eligibility: $\geq$140 mmHg (5 hypertensive bins)
  \item GBD RRs anchor bin-level baseline incidence
  \item \alert{Diabetes subgroup: base control rates assumed 15 percentage points higher than general HTN}
  \item \alert{Diabetes subgroup: multiplicative reduction blended with general HTN effects using diabetes prevalence and high blood pressure prevalence as weights}
\end{itemize}

\medskip
\textbf{Combination of interventions}
\begin{itemize}\footnotesize
  \item Effects combine \textbf{multiplicatively} across all interventions
  \item BP-related effects combined first, then with statins
  \item No saturation, synergy, or interaction terms beyond multiplicative combination
\end{itemize}
\end{column}
\end{columns}

<!-- ======================================================================
     RESULTS
     ====================================================================== -->

# Results

## Baseline Disease Trends (Business-as-Usual)

```{r fig-baseline-trend, fig.height=3.4}
knitr::include_graphics(png_trend)
```

\footnotesize Incident and prevalent CVD cases 2019--2050; solid = calibration, dashed = projection.

## Age-Standardised CVD Mortality Rate

```{r fig-asmr, fig.height=3.0}
ggplot(sl_dt_asmr,
       aes(x = year, y = ASMR, color = int_lbl, linetype = year > 2019)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Business-as-Usual" = "#999999", int_cols),
                     name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"),
                        guide = "none") +
  labs(x = "Year", y = "ASMR (per 100,000)",
       subtitle = "Solid = calibration; dashed = projection") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        panel.grid.minor = element_blank(),
        plot.subtitle = element_text(color = "grey40", size = 8)) +
  guides(color = guide_legend(nrow = 2))
```

## Summary: Cumulative Deaths Averted, 2026--2050

```{r tab-summary, results='asis'}
kable(sl_tab_summary[, .(
        Scenario           = intervention,
        `Cumulative (M)`   = cum_M,
        `Avg./year (M)`    = avg_yr,
        `% of BAU deaths`  = pct_s)],
      format = "latex", booktabs = TRUE, caption = NULL) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9) |>
  row_spec(which(sl_tab_summary$intervention == "All Interventions"),
           bold = TRUE, color = "white", background = "#6a3d9a")
```

\begin{exampleblock}{Headline}
\textbf{All Interventions}: \textbf{`r if(!is.na(total_all)) paste0(fmt_M(total_all)," million deaths averted") else "[X.X M]"`} globally, 2026--2050 ($\approx$ `r if(!is.na(total_all)) paste0(fmt_M(total_all/25)," M/year") else "[X.X M/yr]"`).
\end{exampleblock}

## Annual Deaths Averted Over Time

```{r fig-deaths-averted, fig.height=3.0}
# ggplot(sl_dt_annual,
#        aes(x = year, y = deaths_delayed / 1e6, fill = intervention)) +
#   geom_area(alpha = 0.85, position = "identity") +
#   scale_fill_manual(values = int_cols[int_order], name = NULL) +
#   scale_y_continuous(labels = comma) +
#   labs(x = "Year", y = "Deaths averted (millions)") +
#   theme_minimal(base_size = 10) +
#   theme(legend.position = "bottom",
#         legend.text = element_text(size = 7),
#         panel.grid.minor = element_blank()) +
#   guides(fill = guide_legend(nrow = 2))
#knitr::include_graphics(png_averted)

ggplot(sl_dt_annual[!intervention %in% c("All Interventions", "b.a.u")],
            aes(x = year, y = deaths_delayed / 1e6,
                color = intervention, fill = intervention)) +
  geom_area(position = "stack", alpha = 0.7, color = "grey20", linewidth = 0.15) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_fill_manual(values  = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "",
    x     = "Year",
    y     = "Deaths averted (millions)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
```

\footnotesize Cumulative deaths averted per intervention relative to business-as-usual.

## Cumulative Deaths Averted Over Time

```{r fig-cum-averted, fig.height=3.0}
#knitr::include_graphics(png_cum_averted)
ggplot(sl_dt_annual[!intervention %in% c("All Interventions", "b.a.u")],
            aes(x = year, y = cum_deaths_delayed / 1e6, color = intervention)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "",
    x     = "Year",
    y     = "Cumulative deaths averted (millions)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
```

\footnotesize Cumulative deaths averted per intervention relative to business-as-usual.

## Deaths Averted by Cause of Death

```{r fig-by-cause, fig.height=3.0}
ggplot(sl_dt_cause,
       aes(x = cause_lbl, y = averted / 1e6, fill = intervention)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = int_cols[int_order], name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, y = "Deaths averted (millions)") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 2))
```

## Deaths Averted by Age Group

```{r fig-by-age, fig.height=3.0}
ggplot(sl_dt_age,
       aes(x = age_group, y = averted / 1e6, fill = intervention)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = int_cols[int_order], name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(x = "Age group", y = "Deaths averted (millions)") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(nrow = 2))
```

<!-- ## Top 15 Countries by Projected Impact -->

<!-- ```{r fig-top-countries, fig.height=3.0} -->
<!-- sl_t15 <- copy(sl_tab_top15) -->
<!-- sl_t15[, location := factor(location, levels = rev(location))] -->
<!-- ggplot(sl_t15, aes(x = averted / 1e6, y = location, fill = region)) + -->
<!--   geom_col(width = 0.7) + -->
<!--   scale_x_continuous(labels = comma, expand = c(0, 0)) + -->
<!--   labs(x = "Deaths averted (millions)", y = NULL, fill = "Region") + -->
<!--   theme_minimal(base_size = 10) + -->
<!--   theme(legend.position = "right", -->
<!--         panel.grid.minor = element_blank(), -->
<!--         panel.grid.major.y = element_blank(), -->
<!--         axis.text.y = element_text(size = 8)) -->
<!-- ``` -->

## Regional Summary

```{r tab-region, results='asis'}
kable(sl_tab_region[, .(
        Region           = region,
        `BAU deaths (M)` = fmt_M(bau),
        `Averted (M)`    = fmt_M(averted),
        `% averted`      = paste0(round(pct, 1), "%"))],
      format = "latex", booktabs = TRUE, caption = NULL) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9) |>
  row_spec(0, bold = TRUE)
```

\vspace{4pt}
\begin{block}{Observation}
Absolute gains largest in high-burden regions; proportional gains reflect current coverage gaps.
\end{block}

<!-- ======================================================================
     DISCUSSION
     ====================================================================== -->

# Discussion

## Interpretation

\begin{columns}[T]
\begin{column}{0.55\textwidth}
\textbf{Key findings}
\begin{itemize}
  \item \textbf{All Interventions} yields largest gains: HTN control + statins are complementary
  \item \textbf{IHD} and \textbf{ischemic stroke} drive the majority of avertable deaths
  \item Gains \textbf{accelerate after 2030} as coverage compounds
  \item \textbf{Older adults (70+)} benefit most in absolute terms
  \item \textbf{Top 15 countries} account for $>$75\% of avertable burden
\end{itemize}
\end{column}
\begin{column}{0.43\textwidth}
\begin{exampleblock}{Policy Implication}
Combined WHO Targets 1 + 2 are synergistic --- neither alone captures the full benefit.
\end{exampleblock}
\vspace{6pt}
\begin{alertblock}{Uncertainty \textcolor{red!70!black}{$\bigtriangleup$}}
Sensitivity analysis to diabetes targeting assumptions and secular trends pending.
\end{alertblock}
\end{column}
\end{columns}

## Limitations

\begin{columns}[T]
% ---------------- Left column ----------------
\begin{column}{0.48\textwidth}

\textbf{Model structure \& assumptions}
\begin{itemize}\footnotesize
  \item Stable transition rates assumed; no dynamic feedback between states
  \item Static BP distribution (log-normal): no secular trend and no endogenous updating as coverage increases
  \item No heterogeneity in adherence: newly HTN covered individuals assumed fully adherent
  \item COVID-19 excess mortality only modelled for 2020--2021; long-term effects excluded
\end{itemize}

\medskip
\textbf{Data \& parameters}
\begin{itemize}\footnotesize
  \item RCT-based effect sizes (e.g., Ettehad $\rho_{CF}$): real-world effectiveness at scale may be lower
  \item GBD 2023 uncertainty intervals not propagated
  \item Baseline $C_0$ from NCD-RisC; measurement error unquantified
  \item BP distributions rely on GBD modelling (not direct surveys in all countries)
  \item Statin coverage sparse in some LMICs; imputed from sub-regional estimates
\end{itemize}

\end{column}

% ---------------- Right column ----------------
\begin{column}{0.48\textwidth}

\textbf{Scenarios \& implementation realism}
\begin{itemize}\footnotesize
  \item Linear scale-up assumed (2026--2030); real programmes may follow non-linear trajectories
  \item No explicit health-system capacity constraints modelled
  \item Interactions with statins, diabetes care, and \alert{BP device availability} not included
\end{itemize}

\end{column}
\end{columns}

## Next Steps

\begin{columns}[T]
\begin{column}{0.55\textwidth}
\textbf{Analysis priorities}
\begin{enumerate}
  \item Sensitivity analysis on effect sizes, trends and coverage
  \item Add life-years gained (LYG) and DALYs as co-primary outcome
  \item Stratify by income group (World Bank)
  \item Conduct Aim 2: simulate alternative scale-up trajectories and implementation scenarios for hypertension control
\end{enumerate}
\end{column}
\begin{column}{0.43\textwidth}
\begin{alertblock}{Pending Decisions \textcolor{red!70!black}{$\bigtriangleup$}}
\begin{itemize}
  \item Journal target?
\end{itemize}
\end{alertblock}
\end{column}
\end{columns}

## Summary

\begin{exampleblock}{\textbf{1. The gains are large and achievable}}
All-interventions: \textbf{`r if(!is.na(total_all)) paste0(fmt_M(total_all)," million deaths averted") else "[X.X million]"`} globally, 2026--2050.
\end{exampleblock}

\begin{block}{\textbf{2. HTN control + statins are synergistic}}
Combined deployment outperforms the sum of either alone.
\end{block}

\begin{block}{\textbf{3. Regional patterns reflect burden and coverage gaps}}
Absolute gains largest in high-burden regions; proportional gains reflect current coverage gaps.
\end{block}

\vspace{6pt}
\begin{center}
\textit{Meeting WHO 2030 CVD targets is not just aspirational --- it is quantifiably lifesaving.}
\end{center}

<!-- ======================================================================
     APPENDIX
     ====================================================================== -->

# Appendix

## Notation: Core Model

\begin{center}{\small
\begin{tabular}{ll}
\toprule
Symbol & Description \\
\midrule
\multicolumn{2}{l}{\textit{Core state-transition model}} \\
$a, s, c, t, \ell$ & Age, sex, cause, year, location \\
$W, K, D$ & Well, Sick, Dead populations \\
$IR_{a,s,c}$, $CF_{a,s,c}$ & Baseline incidence \& case-fatality rates \\
$\varepsilon_{IR,j}$, $\varepsilon_{CF,j}$ & Multiplicative effect ratios for intervention $j$ \\
$BG.mx$, $BG.mx.all$ & Cause-specific \& all-cause background mortality \\
$covid.mx$ & COVID-19 mortality (2020--2021 only) \\
\bottomrule
\end{tabular}
}\end{center}

## Notation: HTN Control

\begin{center}{\small
\begin{tabular}{ll}
\toprule
Symbol & Description \\
\multicolumn{2}{l}{\textit{HTN control (general \& diabetes subgroup)}} \\
$C_0$, $C_t$, $C_{\text{target}}$ & Baseline, time-$t$, and target control rates \\
$\Delta C_{\text{target}}$, $\Delta C_{\text{increment}}$ & Total \& time-varying incremental coverage \\
$C_{\text{agg}}^{\Delta}$ & Population-weighted incremental coverage (hypertensive bins) \\
$P(\text{BP}_{\text{cat}})$ & BP category population distribution (8 bins) \\
$RR_{\text{GBD}}$ & GBD relative risk per 10 mmHg SBP increase \\
$\alpha_{\text{inc}}$ & Incidence normalization factor \\
$E_{\text{trial}}$, $E_{\text{DM}}$, $E_{\text{noDM}}$ & Ettehad effect sizes (blended, DM-only, non-DM) \\
$w_{\text{DM}}$ & Diabetes prevalence weight (general HTN: 0.10; DM subgroup: 1.0) \\
$\rho_{CF,c}$ & CF reduction factor per unit incremental coverage \\
$T_{\text{start}}$, $T_{\text{end}}$ & Scale-up start (2026) \& end (2030) \\
\bottomrule
\end{tabular}
}\end{center}

## Notation: Statins \& Combined Effects

\begin{center}{\small
\begin{tabular}{ll}
\toprule
Symbol & Description \\
\midrule
\multicolumn{2}{l}{\textit{Statin therapy}} \\
$U_{0,\ell}$, $U_{\text{target}}$ & Baseline \& target statin coverage \\
$\Delta U(t,\ell)$ & Incremental statin coverage above $U_{0,\ell}$ at time $t$ \\
$T_{\text{stat,start}}$, $T_{\text{stat,end}}$ & Statin scale-up start (2026) \& end years \\
$RR^{IR}_c$, $RR^{CF}_c$ & Trial-based RRs for incidence \& case fatality \\
$\delta^{IR}_c$, $\delta^{CF}_c$ & Risk reductions $= 1 - RR$ \\
$AF_{c,\ell}$ & Cause- \& location-specific attributable fraction (incidence only) \\
$\psi_{\text{athero}}$ & Atherosclerotic fraction of ischemic stroke ($= 0.60$) \\
$E^{IR}_{c,\ell}(t)$, $E^{CF}_{c,\ell}(t)$ & Incremental effect sizes for $IR$ and $CF$ \\
\bottomrule
\end{tabular}
}\end{center}

## HTN Intervention Module: 9-Step Pipeline

```{r steps-table}
steps <- data.table(
  Step=paste0(1:9),
  Action=c(
    "Baseline BP distribution $P(\\text{BP}_{\\text{cat}})$ without treatment (rx$\\,=\\,0$)",
    "GBD relative risks per 10 mmHg SBP $\\to$ bin-specific incidence $IR_{\\text{bin}}$",
    "Diabetes-weighted Ettehad trial effect sizes $E_{\\text{trial}}$ per BP bin $\\times$ cause",
    "Linear coverage scale-up from $C_0$ to $C_{\\text{target}}$, \\textbf{starting 2026}",
    "Coverage adjustment: $E_{\\text{adj}} = E_{\\text{trial}}(C_t - C_0)\\,/\\,(1 - E_{\\text{trial}}\\,C_0)$",
    "New bin incidence: $IR_{\\text{bin,new}} = IR_{\\text{bin}} \\times (1 - E_{\\text{adj}})$",
    "Aggregate: $IR_{\\text{new}} = \\sum IR_{\\text{bin,new}} \\times P(\\text{BP}_{\\text{cat}})$",
    "Incidence effect ratio: $\\varepsilon_{IR} = IR_{\\text{new}}\\,/\\,IR_{\\text{original}}$",
    "Case-fatality effect ratio: $\\varepsilon_{CF} = 1 - \\rho_{CF} \\times C_{\\text{agg}}^{\\Delta}$"
  ),
  Output=c(
    "$P(\\text{BP}_{\\text{cat}}\\,|\\,a,s,t)$",
    "$IR_{\\text{bin}}$","$E_{\\text{trial}}$",
    "$C_t(t,\\,\\text{BP}_{\\text{cat}})$","$E_{\\text{adj}}(t)$",
    "$IR_{\\text{bin,new}}$","$IR_{\\text{new}}$",
    "$\\varepsilon_{IR}\\to$ model","$\\varepsilon_{CF}\\to$ model"
  )
)
kable(steps, col.names=c("","Action","Output"),
      align="cll", escape=FALSE, booktabs=TRUE, format="latex") |>
  kable_styling(latex_options=c("striped","hold_position"),
                font_size=7.5, full_width=TRUE) |>
  column_spec(1, bold=TRUE, width="0.5cm") |>
  column_spec(2, width="9.5cm") |>
  column_spec(3, italic=TRUE, width="2.8cm") |>
  row_spec(c(8,9), background="#d4edda") |>
  row_spec(4, background="#dbeafe")
```

\small\textit{Blue row = intervention start (2026). Green rows = outputs passed to state-transition model.}

## Key Equations

\begin{columns}[T]
\begin{column}{0.48\textwidth}
\textbf{Incidence pathway}

\medskip
\textit{Step 2 --- GBD RR anchors bin incidence:}
$$IR_{\text{bin}} = \frac{RR_{\text{GBD}} \times IR}{\alpha},\quad \alpha = \sum_{\text{cat}} P_{\text{cat}}\cdot RR_{\text{GBD}}$$

\textit{Step 5 --- baseline-adjusted effect (no double-counting):}
$$E_{\text{adj}}(t) = \frac{E_{\text{trial}}\,(C_t - C_0)}{1 - E_{\text{trial}}\,C_0}$$

\textit{Steps 7--8 --- population-weighted effect ratio:}
$$\varepsilon_{IR} = \frac{\sum IR_{\text{bin,new}} \cdot P_{\text{cat}}}{IR_{\text{original}}}$$
\end{column}

\begin{column}{0.48\textwidth}
\textbf{Case fatality pathway}

\medskip
\textit{Incremental coverage above baseline:}
$$\Delta C_{\delta}(t) = \max(C_t - C_0,\;0)$$

\textit{Aggregate across hypertensive bins ($\geq$140 mmHg):}
$$C_{\text{agg}}^{\Delta} = \frac{\sum_{\geq140}\Delta C_{\delta}\cdot P_{\text{cat}}}{\sum_{\geq140} P_{\text{cat}}}$$

\textit{CF effect ratio:}
$$\varepsilon_{CF} = 1 - \rho_{CF}\times C_{\text{agg}}^{\Delta}$$

\small
\begin{tabular}{@{}lc@{}}
\toprule Cause & $\rho_{CF}$ \\ \midrule
IHD & 0.24 \\ Ischemic stroke & 0.36 \\
Hemorrhagic stroke & 0.76 \\ HHD & 0.20 \\
\bottomrule
\end{tabular}
\end{column}
\end{columns}

## HTN Control (Diabetes Subgroup)

\begin{columns}[T]
\begin{column}{0.52\textwidth}
\textbf{WHO Target:} 80\% BP control among diagnosed diabetics by 2030

\medskip
\textbf{Same 9-step pipeline as general HTN.} Key distinction: diabetes-specific Ettehad effect sizes ($w_{\text{DM}} = 1$):

\medskip
\textit{General HTN (DM-blended):}
$$E_{\text{trial}} = (1-w_{\text{DM}})\,E_{\text{noDM}} + w_{\text{DM}}\,E_{\text{DM}},\quad w_{\text{DM}}=0.10$$

\textit{Diabetes subgroup (DM-only):}
$$E_{\text{trial}}^{\text{DM}} = E_{\text{DM}}$$

\textit{Coverage scale-up to $C_{\text{target}}^{\text{DM}} = 0.80$:}
$$C_t^{\text{DM}} = C_0^{\text{DM}} + \Delta C^{\text{DM}}\;\min\!\left[\frac{t-2026}{4},\,1\right]$$
\end{column}

\begin{column}{0.46\textwidth}
\textbf{Effect ratios} (same structure as general HTN):
$$\varepsilon_{IR,\text{DM}}(a,s,c,t) = \frac{IR^{\text{DM}}_{\text{new}}}{IR_{\text{original}}}$$
$$\varepsilon_{CF,\text{DM}} = 1 - \rho_{CF,c} \times C_{\text{agg}}^{\Delta,\text{DM}}$$

\begin{exampleblock}{Why a separate module?}
\small People with diabetes have higher CVD risk at any BP level. DM-specific Ettehad effects are larger (especially for stroke) than blended effects at $w_{\text{DM}}=0.10$.
\end{exampleblock}

\begin{alertblock}{When combined with general HTN}
\small Effects multiply (no double-counting via incremental coverage formulation):
$$\varepsilon_{IR,\text{BP,total}} = \varepsilon_{IR,\text{gen}} \times \varepsilon_{IR,\text{DM}}$$
\end{alertblock}
\end{column}
\end{columns}

## Statin Therapy: Coverage Scale-Up and Key Equations

\begin{columns}[T]
\begin{column}{0.50\textwidth}
\textbf{Scope:} IHD \& ischemic stroke; ages $\geq 40$ only

\medskip
\textit{Incremental coverage above baseline $U_{0,\ell}$:}
$$\Delta U(t,\ell) = \Delta U_{\text{target},\ell}\;\min\!\left[\frac{t-T_{\text{stat,start}}+1}{T_{\text{stat,end}}-T_{\text{stat,start}}+1},1\right]$$
where $\Delta U_{\text{target},\ell} = \max(U_{\text{target}} - U_{0,\ell},\,0)$

\medskip
\textit{Trial-based RRs (per $\approx$1 mmol/L LDL-C reduction):}

\small
\begin{tabular}{@{}lcc@{}}
\toprule Cause & $RR^{IR}$ & $RR^{CF}$ \\ \midrule
IHD & 0.74 & 0.80 \\
Ischemic stroke & 0.80 & 0.96 \\
\bottomrule
\end{tabular}

\normalsize\medskip
\textit{Incidence (primary prevention, AF-scaled):}
$$E^{IR}_{c,\ell}(t) = \frac{AF_{c,\ell}\;\delta^{IR}_{c}\;\Delta U(t,\ell)}{1 - \delta^{IR}_{c}\;U_{0,\ell}}$$
\end{column}

\begin{column}{0.48\textwidth}
\textit{Case fatality (secondary prevention; no AF; atherosclerotic restriction for stroke):}
$$E^{CF}_{c,\ell}(t) = \begin{cases}
\dfrac{\delta^{CF}_{\text{ihd}}\;\Delta U(t,\ell)}{1 - \delta^{CF}_{\text{ihd}}\;U_{0,\ell}}, & c=\text{ihd}\\[1.2em]
\dfrac{\psi_{\text{athero}}\;\delta^{CF}_{\text{ist}}\;\Delta U(t,\ell)}{1 - \delta^{CF}_{\text{ist}}\;U_{0,\ell}}, & c=\text{istroke}
\end{cases}$$

\small $\delta^{IR}_c = 1 - RR^{IR}_c$;\quad $\psi_{\text{athero}} = 0.60$ (atherosclerotic fraction)

\normalsize\medskip
\textit{Effect ratios passed to model:}
$$\varepsilon_{IR,\text{stat}} = \begin{cases} 1 - E^{IR}_{c,\ell}(t), & a\geq40,\;c\in\{\text{ihd,istroke}\} \\ 1, & \text{otherwise} \end{cases}$$
$$\varepsilon_{CF,\text{stat}} = \begin{cases} 1 - E^{CF}_{c,\ell}(t), & a\geq40,\;c\in\{\text{ihd,istroke}\} \\ 1, & \text{otherwise} \end{cases}$$
\end{column}
\end{columns}


## References {.allowframebreaks}

\footnotesize

- \textbf{Ettehad D, et al.}\ Blood pressure lowering for prevention of CVD and death. \textit{Lancet.} 2016;387:957--967.

- **GBD 2023 Collaborators** (2024). Global mortality for 371 causes. *Lancet*.
- **Kontis V et al.** (2019). Contributions of risk factor trends to cardiometabolic mortality decline. *Nature Medicine*, 25, 579--584.
- **Murray CJL et al.** (2020). Global burden of 87 risk factors: GBD 2019. *Lancet*, 396, 1223--1249.


- \textbf{Mills KT, et al.}\ Global disparities of hypertension prevalence and control. \textit{Circulation.} 2016;134:441--450.

- \textbf{NCD Countdown 2030 Collaborators.}\ Pathways to achieving SDG target 3.4. \textit{Lancet.} 2022;399:1226--1249.

- **UNWPP** (2024). World Population Prospects 2024. United Nations.

- **WHO** (2023). Global Hearts Initiative: HEARTS Technical Package. Geneva.

- \textbf{WHO.}\ Global NCD Political Declaration. Geneva: WHO; 2025.


---

\begin{center}
\vfill
{\Large \textbf{Thank You}}\\[10pt]
{\normalsize Questions \& Discussion}\\[16pt]
\begin{beamercolorbox}[sep=6pt,center,rounded=true]{block body example}
\small Contact: David Watkins \textbar{} davidaw@uw.edu \textbar{} University of Washington
\end{beamercolorbox}
\vfill
\end{center}
