---
title: "Projected Health Outcomes of Achieving WHO Targets for Cardiovascular Risk-Factor Control: Probabilistic Sensitivity Analysis"
subtitle: "BP intervention-effect parameter uncertainty (BPLTTC 2021 Appendix S6)"
date: "`r Sys.Date()`"
author: "University of Washington"
params:
  run_dir: ""
  repo_root: ""
  psa_root: ""
  render_token: ""
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

<!--
Render only through code/09_uncertainty_psa.R (psa_render_documents()), which
compiles a private copy outside the source tree and copies the HTML to
output_psa/rendered/. This report reads precomputed PSA artefacts from
params$run_dir; it never reruns the model, the PSA aggregation or the 07/08
calculations. Every file it writes goes through psa_out() (output_psa/ only).
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE,
                      fig.height = 5, fig.width = 9, fig.align = "center")
options(knitr.kable.NA = "—")   # undefined cells (e.g. MC error of a degenerate outcome)
suppressPackageStartupMessages({
  library(data.table); library(ggplot2); library(knitr); library(kableExtra)
  library(DT); library(scales); library(openxlsx); library(sf)
  library(rnaturalearth); library(rnaturalearthdata)
})
if (!identical(params$render_token, "rendered-by-09_uncertainty_psa")) {
  stop("Render through psa_render_documents() in code/09_uncertainty_psa.R so outputs stay in output_psa/.")
}
stopifnot("params$run_dir, repo_root and psa_root are required" =
            nzchar(params$run_dir) && nzchar(params$repo_root) && nzchar(params$psa_root))
repo_root <- normalizePath(params$repo_root, winslash = "/")
psa_root  <- normalizePath(params$psa_root, winslash = "/", mustWork = FALSE)
run_dir   <- normalizePath(params$run_dir, winslash = "/")
# Helper definitions only (path guards, labels, psa_summarise); nothing is recomputed.
for (f in c("091_psa_parameters.R", "092_psa_model.R", "093_psa_downstream.R")) {
  sys.source(file.path(repo_root, "code", f), envir = environment())
}
if (!psa_is_within(run_dir, psa_root)) stop("params$run_dir must be a run inside output_psa/.")
wd_data <- paste0(repo_root, "/data/processed/")
wd_raw  <- paste0(repo_root, "/data/raw/")
fig_dir    <- psa_out(psa_root, "figures")
slides_dir <- psa_out(psa_root, "slides")
paper_dir  <- psa_out(psa_root, "paper")
table_dir  <- psa_out(psa_root, "tables")
for (d in c(fig_dir, slides_dir, paper_dir, table_dir)) dir.create(d, recursive = TRUE, showWarnings = FALSE)
save_rds <- function(obj, dir, file) psa_save_rds(obj, psa_out(dir, file), psa_root)
save_png <- function(plot, dir, file, ...) {
  p <- psa_out(dir, file)
  psa_with_retry(function() ggplot2::ggsave(p, plot = plot, ...))
}
man  <- readRDS(file.path(run_dir, "manifest.rds"))
ep   <- readRDS(file.path(run_dir, "endpoints", "endpoints_draws.rds"))
sm   <- readRDS(file.path(run_dir, "endpoints", "endpoints_summary.rds"))
bau  <- readRDS(file.path(run_dir, "endpoints", "bau_reference_series.rds"))
conv <- readRDS(file.path(run_dir, "endpoints", "convergence.rds"))
reg  <- readRDS(file.path(run_dir, "parameters", "parameter_register.rds"))
countries_dt <- readRDS(file.path(run_dir, "endpoints", "country_metadata_report.rds"))
val_file <- file.path(run_dir, "validation", "validation_results.rds")
val <- if (file.exists(val_file)) readRDS(val_file) else NULL
n_mc <- man$n_mc_draws
is_final <- isTRUE(man$is_final)
if (man$run_mode == "production" && n_mc < man$min_final_draws) {
  stop("Refusing to label ", n_mc, "-draw quantiles as final 95% intervals.")
}
int_lbl  <- if (is_final) "95% PUI" else "TEST interval"
int_long <- if (is_final) {
  sprintf("95%% BP-effect parameter-uncertainty interval (2.5th-97.5th percentiles of %s Monte Carlo draws)",
          format(n_mc, big.mark = ","))
} else {
  sprintf("TEST interval from %d draws (software smoke test; NOT a reportable 95%% interval)", n_mc)
}
```

```{r helpers}
cause_map <- c(ihd = "Ischemic heart disease", istroke = "Ischemic stroke",
               hstroke = "Intracerebral hemorrhage", hhd = "Hypertensive heart disease")
int_labels <- c("b.a.u" = "Business-as-Usual", "HTN Control (No Diabetes)" = "HTN Control (No Diabetes)",
                "HTN Control (Diabetes)" = "HTN Control (Diabetes)",
                "Improved Statin Uptake" = "Improved Statin Uptake", "All Interventions" = "All Interventions")
int_order <- PSA_INT_ORDER
int_cols <- c("HTN Control (No Diabetes)" = "#1f78b4", "HTN Control (Diabetes)" = "#a6cee3",
              "Improved Statin Uptake" = "#e31a1c", "All Interventions" = "#6a3d9a",
              "Business-as-Usual" = "#999999", "HTN Control (Combined, 150M)" = "#33a02c")
cause_cols <- c(ihd = "#56B4E9", istroke = "#CC79A7", hstroke = "#F0E442", hhd = "#E69F00")
region_levels <- c("Africa", "Americas", "Eastern Mediterranean", "Europe",
                   "South-East Asia", "Western Pacific")
income_order <- c("Low income", "Lower-middle income", "Upper-middle income", "High income")
fmt_m <- function(x, digits = 1) formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")
fmt_k <- function(x, digits = 0) formatC(round(x / 1e3, digits), format = "f", digits = digits, big.mark = ",")
fmt_n <- function(x, scale = 1, digits = 0) formatC(round(x / scale, digits), format = "f", digits = digits, big.mark = ",")
DEG <- "‡"
# "point [lower-upper]"; a degenerate interval (identical in every draw) is flagged.
fmt_pi <- function(point, lower, upper, degenerate, scale = 1e6, digits = 2, pct = FALSE) {
  f <- if (pct) function(x) formatC(round(x * 100, digits), format = "f", digits = digits)
       else function(x) fmt_n(x, scale, digits)
  suf <- if (pct) "%" else ""
  ifelse(degenerate, paste0(f(point), suf, " ", DEG),
         paste0(f(point), suf, " [", f(lower), "–", f(upper), suf, "]"))
}
S <- function(tab, var) copy(sm[[tab]][variable == var])
fmt_row <- function(x, ...) fmt_pi(x$point, x$lower, x$upper, x$degenerate, ...)
note_int <- paste0("Point = deterministic model (06 result, draw 0); brackets = ", int_long, ". ",
                   DEG, " = identical in every draw: no sampled parameter enters this estimate ",
                   "(statins and BAU are fixed; All Interventions loses the BP incidence ",
                   "effect through 06 defect D1).")
scen_note <- function() cat(sprintf("\n\n*%s*\n\n", note_int))
theme_psa <- function(base = 13) theme_minimal(base_size = base) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())
cap_int <- paste0("Error bars / ribbons: ", int_long, ".")
```

`r if (!is_final) paste0('<div style="background:#b30000;color:white;padding:12px 16px;font-weight:bold;font-size:1.1em;border-radius:4px;">TEST &mdash; ', n_mc, ' draws; intervals not for reporting. This render is a software smoke test of the PSA pipeline.</div>')`

---

# About this uncertainty analysis

This version of the Aim 1 report adds **probabilistic parameter-uncertainty intervals**
to the estimates of the original report. The analysis is a **probabilistic sensitivity
analysis of BP intervention-effect parameters only**:

* **Sampled:** the `r sum(reg$status == "SAMPLED")` BPLTTC 2021 Appendix S6 hazard ratios per
  5-mmHg SBP reduction that the model's effect workbook actually traverses (ischaemic heart
  disease, stroke and heart failure, each at baseline SBP 130–139, 140–149, 150–159,
  160–169 and ≥170 mmHg). Each is drawn as log(HR) ~ Normal(log HR, SE²) with
  SE = (log U − log L)/3.92 from its reported 95% CI, independently (no covariance is
  reported), and shared by all countries, ages, sexes, years and scenarios within a draw.
  The workbook's `Sheet1` columns are rebuilt from the draws with the workbook's own
  formulas (RR per 10 mmHg = HR5², products over traversed bins).
* **Fixed:** everything else — BP case-fatality factors, all statin parameters, the trans-fat
  (legacy) pathway, baseline epidemiology and demography, BP distributions, diabetes shares,
  policy targets and trajectories, disability weights, life expectancy and all economic
  parameters. The intervals therefore understate total uncertainty and must be read as
  *partial parameter-uncertainty intervals*, not full model uncertainty intervals.
* **Point estimates** are the deterministic model results (draw 0), which reproduce the saved
  06 outputs exactly; **intervals** are the 2.5th and 97.5th percentiles of each estimand
  computed *within* each of the `r format(n_mc, big.mark = ",")` matched Monte Carlo draws
  (baseline and counterfactuals share every draw).
* **Structural assumption carried over unchanged:** Appendix S6 strata are baseline-SBP subgroup
  treatment effects, but the model multiplies them across traversed BP bins as if they were
  successive 10-mmHg increments. Alternative constructions are structural scenarios, not part of
  this PSA.

Run: `r man$run_id` (seed `r man$seed`; mode `r man$run_mode`; `r man$n_countries` countries;
created `r man$created`). Model script: `r paste0("<code>", man$script06, "</code>")` (md5
`r man$script06_md5`); statin adherence arguments in its executed run call:
`r man$run_args_06$adherence_ir` (incidence) / `r man$run_args_06$adherence_cf` (case fatality).

## Estimates without sampled-parameter variation, and inherited 06 defects

The model is reproduced exactly as implemented, including known defects (identified in the
Gate 1 deterministic review); none is corrected here:

* **D1 — All Interventions has no BP incidence effect.** Its legacy sodium step (with zero
  sodium effect) returns missing incidence ratios from 2026, and 06 resets the combined
  incidence ratio to 1. The only sampled parameters therefore never reach this scenario, and
  its intervals are **degenerate (`r DEG`)**. This is a property of the current model, not
  evidence of certainty.
* **Improved Statin Uptake** and the **business-as-usual** series contain no sampled parameter,
  so their intervals are also degenerate.
* **D2** — All Interventions also runs the legacy trans-fat elimination (IHD case fatality).
* **D3** — the scenario ASMR denominator sums `well + sick` over the four cause chains
  (population counted four times); **D4** — the ASMR weights are a decade-step approximation
  labelled as the WHO standard; **D5** — tables sum from 2025 and the % of BAU uses a
  2025–2050 denominator (2025 deaths averted are exactly zero).
* **D9** — `eff_ir/eff_cf` tracking columns differ from applied ratios (not used downstream);
  **D10** — a few initial `well` states are negative (inherited).
* The original executive deck labels its burden-of-disease table columns by position, which
  swaps "DALYs" and "YLLs"; the PSA deck labels them by name.

## Parameter register

```{r register}
samp <- reg[status == "SAMPLED", .(Parameter = param_id, Outcome = outcome, `Baseline SBP` = stratum,
  `HR per 5 mmHg (95% CI)` = sprintf("%.2f (%.2f–%.2f)", estimate, ci_lower, ci_upper),
  meanlog = round(meanlog, 4), sdlog = round(sdlog, 4), `Maps to` = maps_to)]
kable(samp, caption = "Sampled parameters (BPLTTC 2021 Appendix S6).") |>
  kable_styling(bootstrap_options = c("striped", "condensed"), font_size = 12, full_width = TRUE)
oth <- reg[status != "SAMPLED", .N, by = .(Status = status, Family = family)]
kable(oth, caption = paste0("Other register entries (", nrow(reg[status != "SAMPLED"]),
                            " rows; full register with rationale in parameters/parameter_register.csv).")) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), font_size = 12, full_width = FALSE)
```

## Validation against the original model

```{r validation}
if (!is.null(val)) {
  mx <- function(x) { x <- x[is.finite(x)]; if (length(x)) signif(max(x), 3) else NA_real_ }
  vt <- val[, .(Checks = .N, Passed = sum(pass %in% TRUE), Failed = sum(pass %in% FALSE),
                `Informational / not compared` = sum(is.na(pass)),
                `Max abs diff` = mx(max_abs_diff), `Max rel diff` = mx(max_rel_diff)),
            by = .(Group = group, Check = sub(" \\(draw [0-9]+\\)$", "", check))]
  kable(vt, caption = "Validation summary (full table: validation/validation_results.csv).") |>
    kable_styling(bootstrap_options = c("striped", "condensed"), font_size = 11, full_width = TRUE)
} else cat("*Validation was not run for this PSA run.*")
```

```{r validation-stale, results='asis'}
if (!is.null(val) && any(grepl("NOT COMPARED", val$note))) {
  cat(sprintf(paste0("\n\n**Not compared:** %d checks against %s were not run, because those reference ",
                     "files are older than the saved 06 outputs (the original step was not re-run after ",
                     "the model was). They are listed above as informational, not counted as passes.\n\n"),
              sum(grepl("NOT COMPARED", val$note)),
              paste(c(V5_07 = "the 07 DALY output", V6_08 = "the 08 valuation outputs",
                      V7_report = "the original report's artefacts")[man$validation_stale_references],
                    collapse = ", ")))
}
```

## Monte Carlo convergence

```{r convergence}
cf <- conv$final[, .(Outcome = outcome, N = n,
  Point = signif(point, 5), Lower = signif(lower, 5), Upper = signif(upper, 5),
  `Rel. MC error (lower)` = round(rel_mc_error_lower, 3),
  `Rel. MC error (upper)` = round(rel_mc_error_upper, 3),
  Status = fifelse(degenerate, "degenerate (no variation)", fifelse(converged, "meets criterion", "NOT converged")))]
kable(cf, caption = paste0("Relative MC error = half-width of the distribution-free 95% CI of each ",
                           "percentile (order statistics) divided by the interval width; criterion <= ",
                           conv$rel_tol, ". ", if (!is_final) "TEST run: not converged by design." else "")) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), font_size = 11, full_width = TRUE)
```

---

# Background

This report presents modelled projections of cardiovascular disease (CVD) mortality under
WHO-endorsed risk-factor control interventions aligned with four priority global targets.
The central research question is:

> *What would be the projected health impact --- in terms of deaths averted and
> life-years gained --- of achieving WHO targets for hypertension control and
> statin therapy coverage, applied simultaneously across `r man$n_countries` countries
> from 2025 to 2050?*

Four causes of CVD death are modelled jointly: ischemic heart disease (IHD), ischemic
stroke, intracerebral hemorrhage (hemorrhagic stroke), and hypertensive heart disease (HHD).
A discrete-time multi-state transition model (Well -> Sick -> Dead) is calibrated to GBD 2023
data and projected forward to 2050 using UNWPP 2024 population estimates.

---

# WHO Priority Targets

| # | WHO Target | Modelled Intervention |
|---|---|---|
| 1 | 150 million more people with hypertension under control by 2030 | **HTN Control (Combined):** additional control among hypertensives, split into mutually exclusive **no-diabetes** and **diabetes** subgroups |
| 2 | >=50% of eligible individuals receiving lipid-lowering therapy and counselling | **Improved Statin Uptake:** 50% statin coverage among eligible adults by 2030 |
| 3 | 80% blood pressure control among people with diagnosed diabetes | **HTN Control (Diabetes):** BP control among hypertensives *with* diabetes (0.80 aspirational ceiling by 2030) |
| 4 | 80% availability of validated automated blood pressure devices in public and private facilities | Operationally embedded in the HTN Control scenarios; enables Targets 1 and 3 |

Target 1 is modelled as two **mutually exclusive and exhaustive** subgroups of the hypertensive
population (no multiplicative cross-term between them); the 150-million target is allocated by a
single shared proportional scale-up on the **fixed 2025 population**. The control-target tables
below are **fixed policy inputs** (no PSA interval applies).

## HTN Control Targets: 150 Million Additional Controlled, by Diabetes Status and WHO Region

```{r control-summary}
htn_tg <- fread(paste0(wd_data, "htn_control_targets_by_loc.csv"))
htn_tg <- merge(htn_tg, countries_dt[, .(location, region)], by = "location", all.x = TRUE)
n_region_unmapped <- htn_tg[is.na(region), uniqueN(location)]
scenario_lbl <- c(baseline = "Baseline (business-as-usual)", bp_no_diabetes_only = "HTN Control (No Diabetes)",
                  bp_diabetes_only = "HTN Control (Diabetes)", bp_combined = "HTN Control (Combined, 150M target)")
subgroup_lbl <- c(htn_no_diabetes = "Without diabetes", htn_diabetes = "With diabetes", total = "All hypertensives")
agg_control <- function(dt) {
  base <- dt[, .(htn_pop_2025 = sum(htn_population_2025), controlled_2025 = sum(controlled_2025),
                 controlled_2030 = sum(controlled_2030), additional_2030 = sum(additional_controlled_2030)),
             by = .(region, scenario_id, subgroup)]
  tot <- dt[, .(subgroup = "total", htn_pop_2025 = sum(htn_population_2025),
                controlled_2025 = sum(controlled_2025), controlled_2030 = sum(controlled_2030),
                additional_2030 = sum(additional_controlled_2030)), by = .(region, scenario_id)]
  rbind(base, tot)
}
control_summary <- rbind(agg_control(htn_tg[!is.na(region)]), agg_control(copy(htn_tg)[, region := "Global"]))
control_summary[, `:=`(coverage_2025 = controlled_2025 / htn_pop_2025,
                       coverage_2030 = controlled_2030 / htn_pop_2025,
                       pct_increase = fifelse(controlled_2025 > 0, additional_2030 / controlled_2025 * 100, NA_real_),
                       scenario_label = scenario_lbl[scenario_id], subgroup_label = subgroup_lbl[subgroup])]
gcomb <- control_summary[region == "Global" & scenario_id == "bp_combined"]
add_no_diabetes <- gcomb[subgroup == "htn_no_diabetes", additional_2030]
add_diabetes    <- gcomb[subgroup == "htn_diabetes", additional_2030]
add_total       <- gcomb[subgroup == "total", additional_2030]
stopifnot(abs((add_no_diabetes + add_diabetes) - 150e6) <= 1, abs(add_total - 150e6) <= 1)
save_rds(control_summary, slides_dir, "sl_control_summary.rds")
save_rds(list(target_total = 150e6, add_total = add_total, add_no_diabetes = add_no_diabetes,
              add_diabetes = add_diabetes, share_no_diabetes = add_no_diabetes / add_total,
              share_diabetes = add_diabetes / add_total, n_region_unmapped = n_region_unmapped),
         slides_dir, "sl_control_scalars.rds")
control_disp <- function(dt, group_col, group_name) {
  d <- copy(dt); setnames(d, group_col, "Group")
  out <- d[, .(Group, `Target population 2025` = comma(htn_pop_2025, accuracy = 1),
               `Controlled 2025 (n)` = comma(controlled_2025, accuracy = 1),
               `Coverage 2025` = percent(coverage_2025, accuracy = 0.1),
               `Controlled 2030 (n)` = comma(controlled_2030, accuracy = 1),
               `Coverage 2030` = percent(coverage_2030, accuracy = 0.1),
               `Additional controlled 2030` = comma(additional_2030, accuracy = 1),
               `% increase vs baseline` = ifelse(is.na(pct_increase), "—", percent(pct_increase / 100, accuracy = 0.1)))]
  setnames(out, "Group", group_name); out[]
}
```

### Global summary, by scenario (all hypertensives)

```{r control-summary-global}
glob_by_scen <- control_summary[region == "Global" & subgroup == "total"]
glob_by_scen[, scenario_label := factor(scenario_label, levels = scenario_lbl)]
setorder(glob_by_scen, scenario_label)
kable(control_disp(glob_by_scen, "scenario_label", "Scenario"), align = c("l", rep("r", 7)),
      caption = "HTN control targets, global totals by scenario (fixed 2025 population; fixed policy input).") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 13) |>
  row_spec(0, bold = TRUE) |> row_spec(which(glob_by_scen$scenario_id == "bp_combined"), bold = TRUE, background = "#eef3fb")
```

### The 150 million split: diabetes vs. no diabetes (Combined scenario, global)

```{r control-summary-split}
split_tab <- control_summary[region == "Global" & scenario_id == "bp_combined"]
split_tab[, subgroup_label := factor(subgroup_label, levels = c("Without diabetes", "With diabetes", "All hypertensives"))]
setorder(split_tab, subgroup_label)
kable(control_disp(split_tab, "subgroup_label", "Subgroup"), align = c("l", rep("r", 7)),
      caption = "The 150-million target split between mutually exclusive subgroups (fixed policy input).") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 13) |>
  row_spec(0, bold = TRUE) |> row_spec(which(split_tab$subgroup == "total"), bold = TRUE, background = "#eef3fb")
```

Of the **`r comma(add_total, accuracy = 1)`** additional people with hypertension controlled by 2030,
**`r comma(add_no_diabetes, accuracy = 1)`** (`r percent(add_no_diabetes / add_total, accuracy = 0.1)`) are
**without diabetes** and **`r comma(add_diabetes, accuracy = 1)`** (`r percent(add_diabetes / add_total, accuracy = 0.1)`)
are **with diabetes**.

### Combined scenario (150M target), by WHO region

```{r control-summary-region}
region_order <- c(region_levels, "Global")
reg_comb <- control_summary[scenario_id == "bp_combined" & subgroup == "total"]
reg_comb[, region := factor(region, levels = region_order)]
setorder(reg_comb, region)
kable(control_disp(reg_comb, "region", "WHO region"), align = c("l", rep("r", 7)),
      caption = "HTN Control (Combined, 150M target) by WHO region (fixed policy input).") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 13) |>
  row_spec(0, bold = TRUE) |> row_spec(which(reg_comb$region == "Global"), bold = TRUE, background = "#eef3fb")
```

`r if (n_region_unmapped > 0) paste0("*Note: ", n_region_unmapped, " location(s) could not be mapped to a WHO region and are included in the Global total only.*") else ""`

### Combined scenario (150M target), by WHO region **and** diabetes subgroup

```{r control-summary-region-subgroup}
rs <- control_summary[scenario_id == "bp_combined" & region %in% region_order]
rs[, region := factor(region, levels = region_order)]
rs[, subgroup := factor(subgroup, levels = c("htn_no_diabetes", "htn_diabetes", "total"))]
setorder(rs, region, subgroup)
kable(rs[, .(`WHO region` = as.character(region), Subgroup = subgroup_lbl[as.character(subgroup)],
             `Target population 2025` = comma(htn_pop_2025, accuracy = 1),
             `Coverage 2025` = percent(coverage_2025, accuracy = 0.1),
             `Coverage 2030` = percent(coverage_2030, accuracy = 0.1),
             `Additional controlled 2030` = comma(additional_2030, accuracy = 1),
             `% increase vs baseline` = percent(pct_increase / 100, accuracy = 0.1))],
      align = c("l", "l", rep("r", 5)), caption = "Combined 150M target by region and subgroup (fixed policy input).") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) |>
  row_spec(0, bold = TRUE) |> collapse_rows(columns = 1, valign = "top")
```

```{r control-summary-region-fig, fig.width=11, fig.height=5.5}
fig_dat <- control_summary[scenario_id == "bp_combined" & subgroup %in% c("htn_no_diabetes", "htn_diabetes") & region != "Global"]
fig_dat[, subgroup_lbl := factor(fifelse(subgroup == "htn_no_diabetes", "Without diabetes", "With diabetes"),
                                 levels = c("Without diabetes", "With diabetes"))]
reg_ord <- fig_dat[, .(tot = sum(additional_2030)), by = region][order(tot), region]
fig_dat[, region := factor(region, levels = reg_ord)]
reg_tot <- fig_dat[, .(tot = sum(additional_2030) / 1e6), by = region]
g <- ggplot(fig_dat, aes(x = additional_2030 / 1e6, y = region, fill = subgroup_lbl)) +
  geom_col(width = 0.7) +
  geom_text(data = reg_tot, aes(x = tot, y = region, label = sprintf("%.1f", tot)), inherit.aes = FALSE, hjust = -0.15, size = 3.6) +
  scale_fill_manual(values = c("Without diabetes" = "#1f78b4", "With diabetes" = "#a6cee3"), name = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.10))) +
  labs(title = "150 Million Additional BP-Controlled by 2030\nby WHO Region and Diabetes Status",
       subtitle = paste0("Fixed 2025 population | global total ", fmt_m(add_total), " M"),
       x = "Additional people controlled (millions)", y = NULL) +
  theme_psa() + theme(legend.position = "bottom")
print(g)
save_png(g, fig_dir, "aim1_psa_control_region_split.png", width = 11, height = 5.5, dpi = 300)
```

---

# Policy Scenarios

## WHO Target 1 --- Hypertension Control (General Population)

```{r policy-bp, fig.width=10, fig.height=5}
covfxn <- tryCatch(fread(paste0(wd_data, "covfxn2.csv")), error = function(e) NULL)
if (!is.null(covfxn)) {
  g <- ggplot(covfxn[, .(Country, Year, scenario_value = Progress)], aes(x = Year, y = scenario_value * 100, group = Country)) +
    geom_line(color = "grey70", linewidth = 0.6, alpha = 0.7) +
    geom_hline(yintercept = 50, linetype = "dashed", color = "#1f78b4", linewidth = 1) +
    annotate("text", x = 2051, y = 51.5, hjust = 1, size = 3.5, label = "50% target (2030)", color = "#1f78b4") +
    scale_x_continuous(breaks = seq(2020, 2050, 5), limits = c(2020, 2051), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
    labs(title = "WHO Target 1 — HTN Control Scale-up by Country",
         subtitle = "Linear scale-up to ≥50% population BP control by 2030 (fixed input)",
         x = "Year", y = "Hypertension Control Rate (%)") + theme_psa()
  print(g)
  save_png(g, fig_dir, "aim1_psa_policy_htn_general.png", width = 11, height = 5, dpi = 300)
}
```

## WHO Target 3 --- BP Control in People with Diabetes

```{r policy-bp-diabetes}
# As in the original report: country-level diabetes BP plot pending a data file.
```

## WHO Target 2 --- Statin Therapy Scale-up

```{r policy-statins}
# The original report reads data/processed/Statins/statin_data.rds, which does not
# exist, so it skips this plot; the same path is used here to keep the coverage.
dt_statin_scen <- tryCatch(readRDS(paste0(wd_data, "Statins/statin_data.rds")), error = function(e) NULL)
if (is.null(dt_statin_scen)) cat("*Statin scale-up plot skipped, as in the original report (input file not found at data/processed/Statins/statin_data.rds).*")
```

---

# Disease Trends (Baseline)

The business-as-usual (BAU) series contain no sampled parameter; they are **fixed reference series**
(identical in every draw) and are shown without intervals.

## Incidence and Prevalence Over Time

```{r disease-trends, fig.width=10, fig.height=5}
dt_melt <- melt(bau$trend[, .(year, new_cases, sick)], id.vars = "year", variable.name = "metric", value.name = "count")
dt_melt[, metric_lbl := fifelse(metric == "new_cases", "Incident cases", "Prevalent cases")]
g <- ggplot(dt_melt, aes(x = year, y = count / 1e6, color = metric_lbl, linetype = year > 2019)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Incident cases" = "steelblue", "Prevalent cases" = "tomato"), name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  scale_y_continuous(labels = comma) +
  labs(title = "Global Incident and Prevalent CVD Cases — Business-as-Usual",
       subtitle = "All four causes combined; solid = observed, dashed = projected (fixed reference series)",
       x = "Year", y = "Persons (millions)", caption = "Source: Model output; ages 20–95+") +
  theme_psa() + theme(legend.position = "bottom")
print(g)
save_png(g, fig_dir, "aim1_psa_disease_trend.png", width = 11, height = 5, dpi = 300)
```

## Age-Standardised Mortality Rate by Cause

```{r asmr-cause, fig.width=11, fig.height=6}
ac <- copy(bau$asmr_cause)[, cause_lbl := cause_map[cause]]
g <- ggplot(ac, aes(x = year, y = index, color = cause_lbl, linetype = year > 2019)) +
  geom_line(linewidth = 1) + geom_hline(yintercept = 100, linetype = "dotted", color = "grey50") +
  scale_color_manual(values = setNames(cause_cols, cause_map[names(cause_cols)]), name = "Cause of death") +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  labs(title = "Age-Standardised CVD Mortality Rate Index — Business-as-Usual",
       subtitle = "Index: 2019 = 100; solid = observed, dashed = projected (fixed reference series)",
       x = "Year", y = "ASMR Index (2019 = 100)",
       caption = "Weights as in the original report (decade-step approximation; review note D4), ages 20–95+") +
  theme_psa() + theme(legend.position = "bottom")
print(g)
save_png(g, fig_dir, "aim1_psa_asmr_cause.png", width = 11, height = 5, dpi = 300)
```

---

# Projected Deaths Averted (2025-2050)

```{r compute-deaths-averted}
cum <- S("cumul", "deaths_delayed")[intervention %in% int_order][order(match(intervention, int_order))]
tv <- function(lbl) cum[intervention == lbl]
total_all <- tv("All Interventions"); total_bp <- tv("HTN Control (No Diabetes)")
total_bpd <- tv("HTN Control (Diabetes)"); total_stat <- tv("Improved Statin Uptake")
bau_total_deaths <- S("cumul", "bau_total_deaths")[1, point]
n_countries <- man$n_countries
study_years <- paste0(PSA_REPORT_YEAR0, "–", max(PSA_YEARS))
fm <- function(x) fmt_pi(x$point, x$lower, x$upper, x$degenerate, digits = 2)
```

Between 2025 and 2050, achieving all three intervention targets simultaneously is projected to avert
approximately **`r fm(total_all)` million deaths** globally. HTN control in the general population
(WHO Target 1) would avert `r fm(total_bp)` million deaths; targeting BP control specifically in
people with diabetes (WHO Target 3) would avert `r fm(total_bpd)` million deaths. Scale-up of
lipid-lowering therapy (WHO Target 2) would avert an additional `r fm(total_stat)` million deaths.
Brackets: `r int_long`; `r DEG` = no sampled-parameter variation (see Section 1.1).

## Table 1 --- Cumulative Deaths Averted by Intervention (2025-2050)

```{r table1}
ave <- S("cumul", "deaths_delayed_ave")[intervention %in% int_order]
tab1 <- merge(cum, ave[, .(intervention, a_point = point, a_lower = lower, a_upper = upper, a_deg = degenerate)],
              by = "intervention")
tab1 <- tab1[order(match(intervention, int_order))]
tab1[, who_target := fcase(
  intervention == "HTN Control (No Diabetes)", "Target 1 (no-diabetes subgroup of the 150M split)",
  intervention == "HTN Control (Diabetes)", "Target 3: BP control in people with diabetes (0.80 ceiling); diabetes subgroup of the 150M split",
  intervention == "Improved Statin Uptake", "Target 2: >=50% eligible adults on lipid-lowering therapy by 2030",
  intervention == "All Interventions", "All WHO targets (1+2+3) combined", default = "")]
tab1_fmt <- tab1[, .(Intervention = intervention, `WHO Target` = who_target,
  `Deaths averted (M, 2025–2050)` = fmt_pi(point, lower, upper, degenerate, digits = 2),
  `Average per year (thousands)` = fmt_pi(a_point, a_lower, a_upper, a_deg, scale = 1e3, digits = 1))]
kable(tab1_fmt, align = "llrr", caption = paste0("Table 1. Cumulative deaths averted (2025–2050) by intervention [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 13) |>
  column_spec(1, width = "6cm") |> column_spec(2, width = "8cm") |> row_spec(0, bold = TRUE) |>
  footnote(general = note_int)
dt_excel_tab1 <- copy(tab1_fmt)
```

## Figure 1 --- Deaths Averted by Intervention (Bar Chart)

```{r figure1, fig.width=9, fig.height=5}
f1 <- copy(tab1)[, intervention := factor(intervention, levels = rev(int_order))]
g <- ggplot(f1, aes(x = point / 1e6, y = intervention, fill = as.character(intervention))) +
  geom_col(width = 0.65) +
  geom_errorbarh(aes(xmin = lower / 1e6, xmax = upper / 1e6), height = 0.25, color = "black") +
  scale_fill_manual(values = int_cols, guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), labels = comma) +
  labs(title = "Projected Deaths Averted by Intervention, 2025–2050",
       x = "Deaths averted (millions)", y = NULL, caption = cap_int) + theme_psa() +
  theme(panel.grid.major.y = element_blank(), plot.caption = element_text(size = 8))
print(g)
save_png(g, fig_dir, "aim1_psa_fig1_deaths_intervention.png", width = 9, height = 5, dpi = 300)
```

## Table 2 --- Cumulative Deaths Averted by Intervention and Cause

```{r table2}
t2 <- S("cause", "deaths_delayed")[intervention %in% int_order]
t2 <- t2[!(cause %in% c("hstroke", "hhd") & intervention == "Improved Statin Uptake")]
t2[, cause_lbl := cause_map[cause]]
t2[, cell := fmt_pi(point, lower, upper, degenerate, digits = 2)]
t2w <- dcast(t2, intervention ~ cause_lbl, value.var = "cell", fill = "–")
t2w[, intervention := factor(intervention, levels = int_order)]; setorder(t2w, intervention)
kable(t2w, align = c("l", rep("r", ncol(t2w) - 1)),
      caption = paste0("Table 2. Cumulative deaths averted (millions, 2025–2050) by intervention and cause [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 12) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste("– = intervention does not affect this cause. Statins affect IHD and ischemic stroke only.", note_int))
dt_excel_tab2 <- copy(t2w)
dt_tab2 <- copy(t2)
```

## Figure 2 --- Deaths Averted by Cause

```{r figure2, fig.width=10, fig.height=5}
f2 <- S("cause", "deaths_delayed")[intervention %in% int_order]
f2[, `:=`(intervention = factor(intervention, levels = int_order), cause_lbl = factor(cause_map[cause], levels = unname(cause_map)))]
g <- ggplot(f2, aes(x = point / 1e6, y = intervention, fill = cause_lbl)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_errorbarh(aes(xmin = lower / 1e6, xmax = upper / 1e6), position = position_dodge(width = 0.8),
                 height = 0.25, color = "black") +
  scale_fill_manual(values = setNames(cause_cols, cause_map[names(cause_cols)]), name = "Cause of death") +
  scale_x_continuous(labels = comma) + guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(title = "Deaths Averted by Cause of Death and Intervention, 2025–2050",
       subtitle = "Bars are shown side by side (not stacked) so each has its own interval",
       x = "Deaths averted (millions)", y = NULL, caption = cap_int) +
  theme_psa() + theme(legend.position = "bottom", panel.grid.major.y = element_blank(), plot.caption = element_text(size = 8))
print(g)
save_png(g, fig_dir, "aim1_psa_fig2_deaths_cause.png", width = 10, height = 5, dpi = 300)
```

## Figure 3 --- Time Series: Annual Deaths Averted by Intervention

```{r figure3-annual, fig.width=10, fig.height=5}
ann <- S("annual", "deaths_delayed")[intervention %in% int_order]
ann[, intervention := factor(intervention, levels = int_order)]
g <- ggplot(ann[intervention != "All Interventions"], aes(x = year, y = point / 1e6, color = intervention, fill = intervention)) +
  geom_ribbon(aes(ymin = lower / 1e6, ymax = upper / 1e6), alpha = 0.25, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = int_cols, name = NULL) + scale_fill_manual(values = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(title = "Annual Deaths Averted by Intervention, 2025–2050",
       subtitle = "Lines with intervals (the original report stacks these areas)",
       x = "Year", y = "Deaths averted (millions)", caption = cap_int) +
  theme_psa() + theme(legend.position = "bottom", plot.caption = element_text(size = 8))
print(g)
save_png(g, fig_dir, "aim1_psa_fig3_annual_averted.png", width = 11, height = 5, dpi = 300)
```

## Figure 4 --- Cumulative Deaths Averted Over Time

```{r figure4-cumulative, fig.width=10, fig.height=5}
cm <- S("annual", "cum_deaths_delayed")[intervention %in% int_order]
cm[, intervention := factor(intervention, levels = int_order)]
g <- ggplot(cm, aes(x = year, y = point / 1e6, color = intervention, fill = intervention)) +
  geom_ribbon(aes(ymin = lower / 1e6, ymax = upper / 1e6), alpha = 0.25, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols[int_order], name = NULL) + scale_fill_manual(values = int_cols[int_order], name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cumulative Deaths Averted by Intervention, 2025–2050", x = "Year",
       y = "Cumulative deaths averted (millions)", caption = cap_int) +
  theme_psa() + theme(legend.position = "bottom", plot.caption = element_text(size = 8))
print(g)
save_png(g, fig_dir, "aim1_psa_fig4_cum_averted.png", width = 11, height = 5, dpi = 300)
```

## Figure 4b --- Cumulative Deaths Averted Over Time by Region (All Interventions)

```{r figure4b-cumulative-region, fig.width=10, fig.height=5}
ra <- S("region_annual", "cum_deaths_delayed")[intervention == "All Interventions" & region %in% region_levels]
ra[, region := factor(region, levels = region_levels)]
g <- ggplot(ra, aes(x = year, y = point / 1e6, color = region, fill = region)) +
  geom_ribbon(aes(ymin = lower / 1e6, ymax = upper / 1e6), alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.1) +
  scale_color_brewer(palette = "Dark2", name = NULL) + scale_fill_brewer(palette = "Dark2", name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cumulative Deaths Averted Over Time by Region, 2025–2050",
       subtitle = paste("All Interventions combined", DEG, "(no sampled-parameter variation: defect D1)"),
       x = "Year", y = "Cumulative deaths averted (millions)") +
  theme_psa() + theme(legend.position = "bottom")
print(g)
save_png(g, fig_dir, "aim1_psa_fig7_cum_region.png", width = 11, height = 5, dpi = 300)
```

---

# Outputs by Demographics

## By Sex and Age Group

```{r demo-age-sex, fig.width=10, fig.height=5}
as_ <- S("age_sex", "deaths_delayed")[intervention %in% int_order]
order_lvls <- as_[, .(total = sum(point)), by = intervention][order(total), intervention]
as_[, `:=`(intervention = factor(intervention, levels = order_lvls), age_group = factor(age_group, levels = PSA_AGE_GROUPS))]
g <- ggplot(as_, aes(x = point / 1e6, y = intervention, fill = age_group)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_errorbarh(aes(xmin = lower / 1e6, xmax = upper / 1e6), position = position_dodge(width = 0.8), height = 0.2) +
  facet_wrap(~sex, nrow = 1) + scale_x_continuous(labels = comma) +
  labs(title = "Deaths Averted by Age Group, Sex, and Intervention (2025–2050)",
       x = "Deaths averted (millions)", y = NULL, fill = "Age group", caption = cap_int) +
  theme_psa() + theme(legend.position = "bottom", strip.text = element_text(face = "bold"),
                      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
                      plot.caption = element_text(size = 8))
print(g)
save_png(g, fig_dir, "aim1_psa_fig5_demo_age.png", width = 10, height = 5, dpi = 300)
```

## Table 3 --- Deaths Averted by Age Group and Intervention

```{r table3-age}
ag <- function(v) S("age", v)[intervention %in% int_order]
t3 <- ag("deaths_delayed")[, .(age_group, intervention, dd = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))]
t3 <- ag("deaths_projected")[, .(age_group, intervention, dp = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))][t3, on = .(age_group, intervention)]
t3 <- ag("baseline_deaths")[, .(age_group, intervention, bd = comma(point, accuracy = 1))][t3, on = .(age_group, intervention)]
t3 <- ag("deaths_delayed_ave")[, .(age_group, intervention, da = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))][t3, on = .(age_group, intervention)]
t3 <- ag("deaths_delayed_pct")[, .(age_group, intervention, pc = fmt_pi(point, lower, upper, degenerate, digits = 1, pct = TRUE))][t3, on = .(age_group, intervention)]
t3[, `:=`(intervention = factor(intervention, levels = int_order), age_group = factor(age_group, levels = PSA_AGE_GROUPS))]
setorder(t3, age_group, intervention)
t3_fmt <- t3[, .(`Age group` = age_group, Intervention = intervention, `Projected deaths` = dp,
                 `Baseline deaths` = bd, `Deaths averted` = dd, `Deaths averted/year` = da, `% averted` = pc)]
kable(t3_fmt, align = c("l", "l", "r", "r", "r", "r", "r"),
      caption = paste0("Table 3. Cumulative deaths averted (2025–2050) by age group and intervention [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11) |>
  row_spec(0, bold = TRUE) |> footnote(general = note_int)
dt_excel_tab3 <- copy(t3_fmt)
```

---

# Outputs by Geography

## Table 4 --- Deaths Averted by Country (All Interventions)

```{r table4-country}
cty <- function(v) S("country", v)[intervention == "All Interventions"]
t4 <- cty("deaths_delayed")[, .(region, location, dd_point = point, dd = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))]
t4 <- cty("deaths_projected")[, .(location, dp = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))][t4, on = "location"]
t4 <- cty("baseline_deaths")[, .(location, bd = comma(point, accuracy = 1))][t4, on = "location"]
t4 <- cty("deaths_delayed_ave")[, .(location, da = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))][t4, on = "location"]
t4 <- cty("deaths_delayed_pct")[, .(location, pc = fmt_pi(point, lower, upper, degenerate, digits = 1, pct = TRUE))][t4, on = "location"]
setorder(t4, region, -dd_point)
t4_fmt <- t4[, .(Region = region, Country = location, `Projected deaths` = dp, `Baseline deaths` = bd,
                 `Deaths averted` = dd, `Deaths averted/year` = da, `% averted` = pc)]
datatable(t4_fmt, rownames = FALSE, filter = "top",
          caption = htmltools::tags$caption(style = "caption-side: top; text-align: left;",
            paste0("Table 4. Country-level deaths averted under all interventions, 2025–2050 [", int_lbl, "]. ", note_int)),
          options = list(pageLength = 15, scrollX = TRUE, dom = "lftipr"))
dt_excel_tab4 <- copy(t4_fmt)
```

## Figure 5 --- Top 20 Countries by Deaths Averted

```{r figure5-top20, fig.width=10, fig.height=7}
top <- cty("deaths_delayed")[order(-point)][1:min(20, .N)]
top[, location := factor(location, levels = rev(location))]
g <- ggplot(top, aes(x = point / 1e6, y = location, fill = region)) +
  geom_col(width = 0.7) + geom_errorbarh(aes(xmin = lower / 1e6, xmax = upper / 1e6), height = 0.3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), labels = comma) +
  labs(title = "Top 20 Countries by Projected Deaths Averted\n(All Interventions, 2025–2050)",
       subtitle = paste(DEG, "All Interventions intervals are degenerate (defect D1)"),
       x = "Deaths averted (millions)", y = NULL, fill = "Region") +
  theme_psa() + theme(legend.position = "right", panel.grid.major.y = element_blank())
print(g)
save_png(g, fig_dir, "aim1_psa_fig5_top20_countries.png", width = 10, height = 7, dpi = 300)
```

## Figure 6 --- World Map: Deaths Averted per 100,000 Population

```{r figure6-map, fig.width=11, fig.height=5}
wpp <- tryCatch(as.data.table(readRDS(paste0(wd_data, "PopulationsSingleAge0050.rds"))), error = function(e) NULL)
if (!is.null(wpp)) {
  pop_2019 <- wpp[year_id == 2019 & age >= 20, .(pop = sum(Nx, na.rm = TRUE)), by = location]
  mp <- merge(ep$country[intervention == "All Interventions", .(draw_id, location, deaths_delayed)], pop_2019, by = "location", all.x = TRUE)
  mp[, da_per100k := deaths_delayed / pop * 1e5]
  map_sum <- psa_summarise(mp, "location", "da_per100k")
  world <- ne_countries(scale = "medium", returnclass = "sf")
  world_dt <- merge(world, merge(map_sum[, .(location, da_per100k = point)], countries_dt[, .(location, iso3)], by = "location", all.x = TRUE),
                    by.x = "adm0_a3", by.y = "iso3", all.x = TRUE)
  g <- ggplot(world_dt) + geom_sf(aes(fill = da_per100k), color = "grey80", linewidth = 0.1) +
    scale_fill_viridis_c(name = "per 100,000", na.value = "grey90", option = "C", direction = -1) +
    labs(title = "Deaths Averted per 100,000 Population, 2025–2050",
         subtitle = paste("All interventions combined (point estimates;", DEG, "no sampled-parameter variation, defect D1)"),
         caption = "Denominator = UNWPP 2024 age 20+ population in 2019") +
    theme_void(base_size = 12) + theme(legend.position = "right", plot.title = element_text(face = "bold"))
  print(g)
  save_png(g, fig_dir, "aim1_psa_fig6_map.png", width = 11, height = 5, dpi = 300)
}
```

## Table 5 --- Deaths Averted by Region (All Interventions)

```{r table5-region}
rg <- function(v) S("region", v)[intervention == "All Interventions"]
t5 <- rg("deaths_delayed")[, .(region, dd_point = point, dd = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))]
t5 <- rg("deaths_projected")[, .(region, dp = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))][t5, on = "region"]
t5 <- rg("baseline_deaths")[, .(region, bd = comma(point, accuracy = 1))][t5, on = "region"]
t5 <- rg("deaths_delayed_ave")[, .(region, da = fmt_pi(point, lower, upper, degenerate, scale = 1, digits = 0))][t5, on = "region"]
t5 <- rg("deaths_delayed_pct")[, .(region, pc = fmt_pi(point, lower, upper, degenerate, digits = 1, pct = TRUE))][t5, on = "region"]
setorder(t5, -dd_point)
t5_fmt <- t5[, .(Region = region, `Projected deaths` = dp, `Baseline deaths` = bd, `Deaths averted` = dd,
                 `Deaths averted/year` = da, `% averted` = pc)]
kable(t5_fmt, align = c("l", "r", "r", "r", "r", "r"),
      caption = paste0("Table 5. Deaths averted by region under all interventions, 2025–2050 [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) |>
  row_spec(0, bold = TRUE) |> footnote(general = note_int)
dt_excel_tab5 <- copy(t5_fmt)
```

---

# ASMR Under Intervention Scenarios

```{r asmr-interventions, fig.width=12, fig.height=7}
asmr_int <- S("asmr", "ASMR")[intervention %in% c("b.a.u", int_order)]
asmr_int[, int_lbl := fifelse(intervention == "b.a.u", "Business-as-Usual", intervention)]
asmr_int[, intervention := factor(intervention, levels = c("b.a.u", int_order))]
g <- ggplot(asmr_int, aes(x = year, y = point, color = int_lbl, fill = int_lbl)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.25, color = NA) +
  geom_line(aes(linetype = year > 2019), linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) + scale_fill_manual(values = int_cols, name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  labs(title = "Age-Standardised CVD Mortality Rate by Intervention Scenario",
       subtitle = "All causes combined; solid = calibration period, dashed = projection",
       x = "Year", y = "ASMR (per 100,000 population)",
       caption = paste("Computed exactly as in the original report (denominator counts population four times, D3; weights D4).", cap_int)) +
  theme_psa() + theme(legend.position = "bottom", plot.caption = element_text(size = 8))
print(g)
save_png(g, fig_dir, "aim1_psa_asmr_interventions.png", width = 12, height = 6, dpi = 300)
```

---

# Burden of Disease Results

Deaths, YLLs, YLDs and DALYs averted (2026--2050) are computed within each draw exactly as in
07 (YLD = prevalent cases x GBD 2023 disability weight; YLL = deaths x WPP 2024 remaining life
expectancy), then summarised.

```{r bod-helpers}
bod_disp <- function(tab, by = NULL) {
  vs <- c("Deaths", "YLLs", "YLDs", "DALYs")
  x <- sm[[tab]][variable %in% vs & intervention %in% int_order]
  x[, cell := fmt_pi(point, lower, upper, degenerate, digits = 2)]
  w <- dcast(x, as.formula(paste(paste(c(by, "intervention"), collapse = " + "), "~ variable")), value.var = "cell")
  setcolorder(w, c(by, "intervention", vs))
  w[]
}
ord_int <- function(w) w[, intervention := factor(intervention, levels = int_order)][]
```

## Table --- Burden of Disease Averted by Intervention Scenario

```{r bod-table1}
b1 <- ord_int(bod_disp("bod")); setorder(b1, intervention); setnames(b1, "intervention", "Intervention")
kable(b1, align = c("l", "r", "r", "r", "r"),
      caption = paste0("Cumulative burden of disease averted (millions, 2026–2050) [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 12) |>
  row_spec(0, bold = TRUE) |> footnote(general = note_int)
```

## Table --- Burden of Disease Averted by Intervention Scenario and Cause

```{r bod-table2}
b2 <- bod_disp("bod_cause", "cause")
pc <- sm$bod_cause[variable %in% c("pct_deaths", "pct_dalys") & intervention %in% int_order]
pc[, cell := fifelse(degenerate, paste0(round(point, 1), "% ", DEG),
                     paste0(round(point, 1), "% [", round(lower, 1), "–", round(upper, 1), "%]"))]
pcw <- dcast(pc, cause + intervention ~ variable, value.var = "cell")
b2 <- merge(b2, pcw, by = c("cause", "intervention"))
b2[, `:=`(Cause = cause_map[cause], intervention = factor(intervention, levels = int_order))]
setorder(b2, intervention, cause)
kable(b2[, .(Intervention = intervention, Cause, Deaths, YLLs, YLDs, DALYs, `% Deaths` = pct_deaths, `% DALYs` = pct_dalys)],
      align = c("l", "l", rep("r", 6)),
      caption = paste0("Burden averted (millions, 2026–2050) by intervention and cause; % = share within scenario, computed within draw [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11) |>
  row_spec(0, bold = TRUE) |> footnote(general = note_int)
```

## Table --- Burden of Disease Averted by WHO Region and Intervention Scenario

```{r bod-table3}
b3 <- ord_int(bod_disp("bod_region", "region")); setorder(b3, region, intervention)
kable(b3[, .(Region = region, Intervention = intervention, Deaths, YLLs, YLDs, DALYs)], align = c("l", "l", rep("r", 4)),
      caption = paste0("Burden averted (millions, 2026–2050) by WHO region and intervention [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11) |>
  row_spec(0, bold = TRUE) |> footnote(general = note_int)
```

## Table --- Burden of Disease Averted by World Bank Income Group and Intervention Scenario

```{r bod-table4}
b4 <- ord_int(bod_disp("bod_income", "location_income"))
b4[, location_income := factor(location_income, levels = income_order)]; setorder(b4, location_income, intervention)
kable(b4[, .(`Income Group` = location_income, Intervention = intervention, Deaths, YLLs, YLDs, DALYs)], align = c("l", "l", rep("r", 4)),
      caption = paste0("Burden averted (millions, 2026–2050) by World Bank income group and intervention [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11) |>
  row_spec(0, bold = TRUE) |> footnote(general = note_int)
```

## Figure --- DALYs Averted Time Series by Intervention Scenario

```{r bod-fig1-dalys-ts, fig.width=10, fig.height=5}
ba <- S("bod_annual", "cum_daly_averted")[intervention %in% int_order]
ba[, intervention := factor(intervention, levels = int_order)]
g <- ggplot(ba, aes(x = year, y = point / 1e6, color = intervention, fill = intervention)) +
  geom_ribbon(aes(ymin = lower / 1e6, ymax = upper / 1e6), alpha = 0.25, color = NA) + geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols[int_order], name = NULL) + scale_fill_manual(values = int_cols[int_order], name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(title = "Cumulative DALYs Averted by Intervention Scenario, 2026–2050", x = "Year",
       y = "Cumulative DALYs averted (millions)", caption = cap_int) +
  theme_psa() + theme(legend.position = "bottom", plot.caption = element_text(size = 8))
print(g)
save_png(g, fig_dir, "aim1_psa_bod_fig1_dalys_ts.png", width = 10, height = 5, dpi = 300)
```

## Figure --- Cumulative DALYs Averted by WHO Region (All Interventions Combined)

```{r bod-fig2-dalys-region, fig.width=10, fig.height=5}
br <- S("bod_region", "DALYs")[intervention == "All Interventions"][order(-point)]
br[, region := factor(region, levels = rev(region))]
g <- ggplot(br, aes(x = point / 1e6, y = region)) + geom_col(fill = "#6a3d9a", width = 0.65) +
  geom_errorbarh(aes(xmin = lower / 1e6, xmax = upper / 1e6), height = 0.25) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), labels = comma) +
  labs(title = "Cumulative DALYs Averted by WHO Region, 2026–2050",
       subtitle = paste("All Interventions combined", DEG, "(no sampled-parameter variation, D1)"),
       x = "DALYs averted (millions)", y = NULL) +
  theme_psa() + theme(panel.grid.major.y = element_blank())
print(g)
save_png(g, fig_dir, "aim1_psa_bod_fig2_dalys_region.png", width = 10, height = 5, dpi = 300)
```

---

# Supplementary: HTN Control (Combined, 150M) and the effect of defect D1

This section is **not in the original report**. `bp_combined` is the 150-million policy scenario;
the original report excludes it from the deaths figures. It is the scenario in which the sampled
BP-effect parameters act most fully, and comparing it with All Interventions shows the size of
the BP incidence pathway that defect D1 removes from the latter.

```{r supp-combined}
sc <- rbind(S("cumul", "deaths_delayed")[, .(intervention, v = "Deaths averted (M, 2025–2050)", point, lower, upper, degenerate)],
            S("bod", "DALYs")[, .(intervention, v = "DALYs averted (M, 2026–2050)", point, lower, upper, degenerate)])
sc <- sc[intervention %in% c("HTN Control (Combined, 150M)", "HTN Control (No Diabetes)", "HTN Control (Diabetes)", "All Interventions")]
sc[, cell := fmt_pi(point, lower, upper, degenerate, digits = 2)]
scw <- dcast(sc, intervention ~ v, value.var = "cell")
kable(scw, caption = paste0("Combined BP scenario vs. components and All Interventions [", int_lbl, "].")) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE, font_size = 12) |>
  footnote(general = note_int)
```

---

# Excel Export

```{r excel-export}
wb <- createWorkbook()
add_sheet <- function(wb, sheet_name, dat, title_str) {
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, title_str, startRow = 1, startCol = 1)
  writeData(wb, sheet_name, paste("Brackets:", int_long), startRow = 2, startCol = 1)
  writeData(wb, sheet_name, dat, startRow = 4, startCol = 1,
            headerStyle = createStyle(textDecoration = "bold", fgFill = "#2E75B6", fontColour = "white", border = "Bottom"))
  setColWidths(wb, sheet_name, cols = seq_len(ncol(dat)), widths = "auto")
}
add_sheet(wb, "T1_Deaths_by_Intervention", as.data.frame(dt_excel_tab1), "Table 1. Cumulative deaths averted (2025-2050) by intervention.")
add_sheet(wb, "T2_Deaths_by_Cause", as.data.frame(dt_excel_tab2), "Table 2. Deaths averted (millions) by intervention and cause.")
add_sheet(wb, "T3_Deaths_by_Age", as.data.frame(dt_excel_tab3), "Table 3. Deaths averted by age group and intervention.")
add_sheet(wb, "T4_Deaths_by_Country", as.data.frame(dt_excel_tab4), "Table 4. Country-level deaths averted (All Interventions, 2025-2050).")
add_sheet(wb, "T5_Deaths_by_Region", as.data.frame(dt_excel_tab5), "Table 5. Deaths averted by region (All Interventions, 2025-2050).")
out_xlsx <- psa_out(table_dir, "aim1_results_tables_psa.xlsx")
psa_with_retry(function() saveWorkbook(wb, out_xlsx, overwrite = TRUE))
```

---

# SLIDE DECK OUTPUTS

All artefacts for `aim1_executive_slides_psa.Rmd` are written to `output_psa/slides/`.

```{r slides-artifacts}
pack <- function(x) x[, .(point, lower, upper, degenerate)]
sc_row <- function(name, x) data.table(name = name, pack(x))
sl_scalars <- list(
  run_id = man$run_id, n_mc_draws = n_mc, is_final = is_final, interval_label = int_lbl,
  interval_long = int_long, n_countries = n_countries, study_years = study_years,
  bau_total_deaths = bau_total_deaths,
  table = rbindlist(list(
    sc_row("total_all", total_all), sc_row("total_bp", total_bp), sc_row("total_bpd", total_bpd),
    sc_row("total_stat", total_stat),
    sc_row("total_dalys_all", S("bod", "DALYs")[intervention == "All Interventions"]),
    sc_row("total_ylls_all", S("bod", "YLLs")[intervention == "All Interventions"]),
    sc_row("total_ylds_all", S("bod", "YLDs")[intervention == "All Interventions"]),
    sc_row("total_dalys_bp", S("bod", "DALYs")[intervention == "HTN Control (No Diabetes)"]),
    sc_row("total_dalys_bpd", S("bod", "DALYs")[intervention == "HTN Control (Diabetes)"]),
    sc_row("total_dalys_stat", S("bod", "DALYs")[intervention == "Improved Statin Uptake"]))))
save_rds(sl_scalars, slides_dir, "sl_scalars.rds")
pct <- S("cumul", "pct_of_bau")[intervention %in% int_order, .(intervention, pct_point = point, pct_lower = lower, pct_upper = upper)]
sl_tab_summary <- merge(tab1[, .(intervention, point, lower, upper, degenerate, a_point, a_lower, a_upper)], pct, by = "intervention")
sl_tab_summary <- sl_tab_summary[order(match(intervention, int_order))]
save_rds(sl_tab_summary, slides_dir, "sl_tab_summary.rds")
sl_tab_region <- merge(rg("deaths_delayed")[, .(region, point, lower, upper, degenerate)],
                       rg("baseline_deaths")[, .(region, bau = point)], by = "region")
sl_tab_region <- merge(sl_tab_region, rg("deaths_delayed_pct")[, .(region, pct_point = point * 100, pct_lower = lower * 100, pct_upper = upper * 100)], by = "region")
setorder(sl_tab_region, -point)
save_rds(sl_tab_region, slides_dir, "sl_tab_region.rds")
sl_tab_top15 <- cty("deaths_delayed")[order(-point)][1:min(15, .N), .(location, region, point, lower, upper, degenerate)]
save_rds(sl_tab_top15, slides_dir, "sl_tab_top15.rds")
sl_dt_cause <- S("cause", "deaths_delayed")[intervention %in% int_order, .(cause, intervention, point, lower, upper, degenerate)]
sl_dt_cause[, cause_lbl := fcase(cause == "ihd", "IHD", cause == "istroke", "Ischemic stroke",
                                 cause == "hstroke", "Hemorrhagic stroke", cause == "hhd", "Hypertensive HD")]
sl_dt_cause[, intervention := factor(intervention, levels = int_order)]
save_rds(sl_dt_cause, slides_dir, "sl_dt_cause.rds")
sl_dt_age <- S("age", "deaths_delayed")[intervention %in% int_order, .(age_group = factor(age_group, levels = PSA_AGE_GROUPS),
                                                                         intervention = factor(intervention, levels = int_order),
                                                                         point, lower, upper, degenerate)]
save_rds(sl_dt_age, slides_dir, "sl_dt_age.rds")
save_rds(asmr_int[, .(year, intervention, int_lbl, point, lower, upper, degenerate)], slides_dir, "sl_dt_asmr.rds")
sl_dt_annual <- merge(ann[, .(intervention, year, point, lower, upper, degenerate)],
                      cm[, .(intervention, year, cum_point = point, cum_lower = lower, cum_upper = upper)], by = c("intervention", "year"))
save_rds(sl_dt_annual, slides_dir, "sl_dt_annual.rds")
bs <- sm$bod[variable %in% c("Deaths", "YLLs", "YLDs", "DALYs") & intervention %in% int_order,
             .(intervention = factor(intervention, levels = int_order), measure = variable, point, lower, upper, degenerate)]
save_rds(bs, slides_dir, "sl_bod_summary.rds")

# Slide PNGs (slide-optimised dimensions, with intervals).
g <- ggplot(dt_melt, aes(x = year, y = count / 1e6, color = metric_lbl, linetype = year > 2019)) + geom_line(linewidth = 1.1) +
  scale_color_manual(values = c("Incident cases" = "steelblue", "Prevalent cases" = "tomato"), name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") + scale_y_continuous(labels = comma) +
  labs(x = "Year", y = "Persons (millions)", subtitle = "Solid = calibration; dashed = projection (fixed reference series)") +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom", panel.grid.minor = element_blank())
save_png(g, slides_dir, "sl_fig_baseline_trend.png", width = 8, height = 4, dpi = 200)
g <- ggplot(asmr_int, aes(x = year, y = point, color = int_lbl, fill = int_lbl)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.25, color = NA) + geom_line(aes(linetype = year > 2019), linewidth = 1) +
  scale_color_manual(values = int_cols, name = NULL) + scale_fill_manual(values = int_cols, name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  labs(x = "Year", y = "ASMR (per 100,000)") + theme_minimal(base_size = 11) +
  theme(legend.position = "bottom", legend.text = element_text(size = 7)) + guides(color = guide_legend(nrow = 2), fill = guide_legend(nrow = 2))
save_png(g, slides_dir, "sl_fig_asmr.png", width = 8, height = 4, dpi = 200)
g <- ggplot(ann[intervention != "All Interventions"], aes(x = year, y = point / 1e6, color = intervention, fill = intervention)) +
  geom_ribbon(aes(ymin = lower / 1e6, ymax = upper / 1e6), alpha = 0.25, color = NA) + geom_line(linewidth = 1) +
  scale_color_manual(values = int_cols, name = NULL) + scale_fill_manual(values = int_cols, name = NULL) +
  labs(x = "Year", y = "Deaths averted (millions)") + theme_minimal(base_size = 11) + theme(legend.position = "bottom", legend.text = element_text(size = 7))
save_png(g, slides_dir, "sl_fig_deaths_averted.png", width = 8, height = 4, dpi = 200)
g <- ggplot(sl_dt_cause, aes(x = cause_lbl, y = point / 1e6, fill = intervention)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_errorbar(aes(ymin = lower / 1e6, ymax = upper / 1e6), position = position_dodge(width = 0.8), width = 0.25) +
  scale_fill_manual(values = int_cols[int_order], name = NULL) + labs(x = NULL, y = "Deaths averted (millions)") +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom", legend.text = element_text(size = 7)) + guides(fill = guide_legend(nrow = 2))
save_png(g, slides_dir, "sl_fig_by_cause.png", width = 8, height = 4, dpi = 200)
g <- ggplot(sl_dt_age, aes(x = age_group, y = point / 1e6, fill = intervention)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_errorbar(aes(ymin = lower / 1e6, ymax = upper / 1e6), position = position_dodge(width = 0.8), width = 0.25) +
  scale_fill_manual(values = int_cols[int_order], name = NULL) + labs(x = "Age group", y = "Deaths averted (millions)") +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom", legend.text = element_text(size = 7)) + guides(fill = guide_legend(nrow = 2))
save_png(g, slides_dir, "sl_fig_by_age.png", width = 8, height = 4, dpi = 200)
t15 <- copy(sl_tab_top15)[, location := factor(location, levels = rev(location))]
g <- ggplot(t15, aes(x = point / 1e6, y = location, fill = region)) + geom_col(width = 0.7) +
  geom_errorbarh(aes(xmin = lower / 1e6, xmax = upper / 1e6), height = 0.3) +
  labs(x = "Deaths averted (millions)", y = NULL, fill = "Region") + theme_minimal(base_size = 10)
save_png(g, slides_dir, "sl_fig_top15_countries.png", width = 8, height = 5, dpi = 200)
g <- ggplot(cm[intervention != "All Interventions"], aes(x = year, y = point / 1e6, color = intervention, fill = intervention)) +
  geom_ribbon(aes(ymin = lower / 1e6, ymax = upper / 1e6), alpha = 0.25, color = NA) + geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) + scale_fill_manual(values = int_cols, name = NULL) +
  labs(x = "Year", y = "Cumulative deaths averted (millions)") + theme_minimal(base_size = 13) + theme(legend.position = "bottom")
save_png(g, slides_dir, "sl_fig_cum_averted.png", width = 11, height = 5, dpi = 300)
```

```{r slides-econ}
# Economic valuation (08 analogue computed within each draw by 09).
econ_scen_lbl <- c(bp_no_diabetes_only = "HTN Control (No Diabetes)", bp_diabetes_only = "HTN Control (Diabetes)",
                   statins_only = "Improved Statin Uptake", all_interventions = "All Interventions")
ec <- sm$econ[scenario %in% names(econ_scen_lbl)]
ec[, intervention := factor(econ_scen_lbl[scenario], levels = int_order)]
build_econ <- function(vt) {
  x <- ec[valuation_type == vt & elasticity_case == "e1_2_primary" & variable %in% c("metric_total", "share_total")]
  dcast(x, who_region + scenario + intervention ~ variable, value.var = c("point", "lower", "upper", "degenerate"))
}
sl_econ_vsl_primary  <- build_econ("VSL");  setorder(sl_econ_vsl_primary, intervention, who_region)
sl_econ_vsly_primary <- build_econ("VSLY"); setorder(sl_econ_vsly_primary, intervention, who_region)
save_rds(sl_econ_vsl_primary, slides_dir, "sl_econ_vsl_primary.rds")
save_rds(sl_econ_vsly_primary, slides_dir, "sl_econ_vsly_primary.rds")
wv <- function(tab, v) tab[who_region == "World" & intervention == "All Interventions"]
sl_econ_scalars <- list(
  world_vsly_share_total = wv(sl_econ_vsly_primary)[, .(point = point_share_total, lower = lower_share_total, upper = upper_share_total, degenerate = degenerate_share_total)],
  world_vsl_share_total  = wv(sl_econ_vsl_primary)[, .(point = point_share_total, lower = lower_share_total, upper = upper_share_total, degenerate = degenerate_share_total)],
  world_vsly_value_total = wv(sl_econ_vsly_primary)[, .(point = point_metric_total, lower = lower_metric_total, upper = upper_metric_total, degenerate = degenerate_metric_total)],
  world_vsl_value_total  = wv(sl_econ_vsl_primary)[, .(point = point_metric_total, lower = lower_metric_total, upper = upper_metric_total, degenerate = degenerate_metric_total)],
  base_year = 2026L, horizon_end = 2050L,
  elasticity_primary = "Differential (e=0.8 HIC / 1.2 LMIC), Robinson et al. 2019", discount_rate = 0.03)
save_rds(sl_econ_scalars, slides_dir, "sl_econ_scalars.rds")
dd <- ec[valuation_type == "VSLY" & grepl("^share_", variable)]
dd[, period := factor(fcase(variable == "share_2026", "2026", variable == "share_2030", "2030", variable == "share_2040", "2040",
                            variable == "share_2050", "2050", variable == "share_total", "2026–2050\n(cumulative)"),
                      levels = c("2026", "2030", "2040", "2050", "2026–2050\n(cumulative)"))]
dd[, elasticity_label := fifelse(elasticity_case == "e1_2_primary", "Primary (e = 0.8/1.2)", "Sensitivity (e = 1.5)")]
sl_econ_dotplot_dt <- dd[, .(valuation_type, elasticity_case, elasticity_label, who_region, scenario, intervention, period,
                             share_income = point, lower, upper, degenerate)]
save_rds(sl_econ_dotplot_dt, slides_dir, "sl_econ_dotplot_dt.rds")
g <- ggplot(sl_econ_dotplot_dt, aes(x = share_income, y = intervention, colour = elasticity_label)) +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0, position = position_dodge(width = 0.55)) +
  geom_point(size = 2.2, alpha = 0.9, position = position_dodge(width = 0.55)) +
  facet_grid(who_region ~ period, scales = "free_x") +
  scale_x_continuous(labels = percent_format(accuracy = 0.1), expand = expansion(mult = c(0.05, 0.15))) +
  scale_colour_manual(name = "Elasticity case", values = c("Primary (e = 0.8/1.2)" = "#2166ac", "Sensitivity (e = 1.5)" = "#d6604d")) +
  labs(title = "Economic value of avoidable CVD mortality, proportion of regional income",
       subtitle = paste0("VSLY only | Discounted at r = 3% (Robinson et al. 2019) | bars: ", int_lbl),
       x = "Proportion of discounted GNI", y = NULL) +
  theme_bw(base_size = 9) +
  theme(strip.text.y = element_text(angle = 0, hjust = 0, face = "bold"), strip.text.x = element_text(face = "bold"),
        strip.background = element_rect(fill = "#f0f0f0", colour = "grey70"), panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(), legend.position = "bottom", plot.title = element_text(face = "bold"),
        axis.text.y = element_text(size = 8))
save_png(g, slides_dir, "sl_econ_dotplot.png", width = 11, height = 6.2, dpi = 220)
```

```{r slides-manifest}
sl_psa_meta <- list(run_id = man$run_id, n_mc_draws = n_mc, is_final = is_final, interval_label = int_lbl,
                    interval_long = int_long, seed = man$seed, parameter_scope = man$parameter_scope,
                    created = format(Sys.time(), "%Y-%m-%d %H:%M"), degenerate_note = note_int,
                    run_mode = man$run_mode, script06 = man$script06,
                    adherence_ir = man$run_args_06$adherence_ir, adherence_cf = man$run_args_06$adherence_cf)
save_rds(sl_psa_meta, slides_dir, "sl_psa_meta.rds")
save_rds(reg[status == "SAMPLED", .(param_id, outcome, stratum, estimate, ci_lower, ci_upper, meanlog, sdlog, maps_to)],
         slides_dir, "sl_psa_params.rds")
sl_manifest <- data.table(
  file = c("sl_psa_meta.rds", "sl_psa_params.rds", "sl_scalars.rds", "sl_tab_summary.rds", "sl_tab_region.rds", "sl_tab_top15.rds",
           "sl_dt_cause.rds", "sl_dt_age.rds", "sl_dt_asmr.rds", "sl_dt_annual.rds", "sl_bod_summary.rds",
           "sl_control_summary.rds", "sl_control_scalars.rds",
           "sl_fig_baseline_trend.png", "sl_fig_asmr.png", "sl_fig_deaths_averted.png", "sl_fig_by_cause.png",
           "sl_fig_by_age.png", "sl_fig_top15_countries.png", "sl_fig_cum_averted.png",
           "sl_econ_vsl_primary.rds", "sl_econ_vsly_primary.rds", "sl_econ_scalars.rds",
           "sl_econ_dotplot_dt.rds", "sl_econ_dotplot.png"),
  consumed_by = c("TEST banner, interval labels", "appendix parameter table", "inline scalars", "tab-summary", "tab-region", "top-15 (not shown, as original)",
                  "fig-by-cause", "fig-by-age", "fig-asmr", "fig-deaths-averted / fig-cum-averted", "BOD slide",
                  "scenario slides", "scenario slides", "fig-baseline-trend", "fig-asmr (alt)", "fig-deaths-averted (alt)",
                  "fig-by-cause (alt)", "fig-by-age (alt)", "top-15 (alt)", "fig-cum-averted (alt)",
                  "econ table", "econ table / headline", "econ headline", "econ dotplot (alt)", "econ dotplot"))
sl_manifest[, `:=`(present = file.exists(file.path(slides_dir, file)), run_id = man$run_id)]
stopifnot("Slide artefacts missing" = all(sl_manifest$present))
save_rds(sl_manifest, slides_dir, "sl_manifest.rds")
kable(sl_manifest, caption = "Slide deck artefact manifest (all files in output_psa/slides/).") |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = TRUE, font_size = 11)
```

---

# PAPER OUTPUTS

```{r paper-outputs}
paper_scalars <- list(
  run_id = man$run_id, interval_label = int_lbl, is_final = is_final, n_countries = n_countries,
  study_years = study_years, bau_total_deaths = bau_total_deaths,
  deaths_averted = cum[, .(intervention, point, lower, upper, degenerate)],
  pct_of_bau = S("cumul", "pct_of_bau")[intervention %in% int_order, .(intervention, point, lower, upper, degenerate)],
  annual_2030_2050 = S("annual", "deaths_delayed")[year %in% c(2030, 2050) & intervention %in% int_order, .(intervention, year, point, lower, upper, degenerate)],
  cause = S("cause", "deaths_delayed")[intervention %in% int_order, .(intervention, cause, point, lower, upper, degenerate)],
  age_all = S("age", "deaths_delayed")[intervention == "All Interventions", .(age_group, point, lower, upper, degenerate)],
  sex_all = S("sex", "deaths_delayed")[intervention == "All Interventions", .(sex, point, lower, upper, degenerate)],
  bod = sm$bod[intervention %in% int_order, .(intervention, measure = variable, point, lower, upper, degenerate)],
  top5 = cty("deaths_delayed")[order(-point)][1:min(5, .N), .(location, point, lower, upper, degenerate)],
  region_tab = rg("deaths_delayed")[, .(region, point, lower, upper, degenerate)],
  income_tab = sm$bod_income[intervention == "All Interventions" & variable %in% c("Deaths", "DALYs"),
                             .(location_income, measure = variable, point, lower, upper, degenerate)])
save_rds(paper_scalars, paper_dir, "paper_scalars.rds")
for (nm in c("cumul", "cause", "region", "country", "annual", "age_sex", "bod", "bod_cause", "bod_region", "bod_income")) {
  save_rds(sm[[nm]], paper_dir, paste0("paper_", nm, "_summary.rds"))
}
```

---

*Report generated on `r Sys.time()` using R `r R.version$major`.`r R.version$minor` from PSA run `r man$run_id`.*
